\newpage
\sstroutine{
   NDF2GASP
}{
   Converts a two-dimensional NDF into a GASP image
}{
   \sstdescription{
      This application converts a two-dimensional NDF into the GAlaxy
      Surface Photometry (GASP) package's format.  See the Notes for the
      details of the conversion.
   }
   \sstusage{
      ndf2gasp in out [fillbad]
   }
   \sstparameters{
      \sstsubsection{
         IN = NDF (Read)
      }{
         The input NDF data structure. The suggested default is the
         current NDF if one exists, otherwise it is the current value.
      }
      \sstsubsection{
         FILLBAD = \_INTEGER (Read)
      }{
         The value used to replace bad pixels in the NDF's data array
         before it is copied to the GASP file.  The value must be in the
         range of signed words (-32768 to 32767).  A null value ({\tt !})
         means no replacements are to be made.  This parameter is
         ignored if there are no bad values.  {\tt [!]}
      }
      \sstsubsection{
         OUT = LITERAL (Write)
      }{
         The name of the output GASP image.  Two files are produced
         with the same name but different extensions.  The {\tt "}.dat{\tt "} file
         contains the data array, and {\tt "}.hdr{\tt "} is the associated header
         file that specifies the dimensions of the image.  The
         suggested default is the current value.
      }
   }
   \sstexamples{
      \sstexamplesubsection{
         ndf2gasp abell1367 a1367
      }{
         Converts an NDF called abell1367 into the GASP image comprising
         the pixel file a1367.dat and the header file a1367.hdr.  If
         there are any bad values present they are copied verbatim to
         the IRAF image.
      }
      \sstexamplesubsection{
         ndf2gasp ngc253 ngc253 fillbad=-1
      }{
         Converts the NDF called ngc253 to a GASP image comprising the
         pixel file ngc253.dat and the header file ngc253.hdr.  Any bad
         values in the data array are replaced by minus one.
      }
   }
   \sstnotes{
      The rules for the conversion are as follows:
      \sstitemlist{

         \sstitem
         The NDF data array is copied to the {\tt "}.dat{\tt "} file.

         \sstitem
         The dimensions of the NDF data array is written to the {\tt "}.hdr{\tt "}
         header file.

         \sstitem
         All other NDF components are ignored.
      }
   }
   \sstdiytopic{
      Related Applications
   }{
      CONVERT: GASP2NDF.
   }
   \sstdiytopic{
      References
   }{
      GASP documentation (MUD/66).
   }
   \sstimplementationstatus{
      \sstitemlist{

         \sstitem
         The GASP image produced has by definition type SIGNED WORD.
         There is type conversion of the data array to this type.

         \sstitem
         Bad values may arise due to type conversion.  These too are
         substituted by the (non-null) value of FILLBAD.

         \sstitem
         For an NDF with an odd number of columns, the last column from
         the GASP image is trimmed.
      }
   }
}
