\newpage
\sstroutine{
   NDF2PGM
}{
   Converts an NDF to a PBMPLUS-style PGM-format file.
}{
   \sstdescription{
      This application converts an NDF to a PBMPLUS PGM-format file.
      The programme first finds the brightest and darkest pixel values 
      in the image.  It then uses these to determine suitable scaling
      factors to convert the image into an 8-bit representation.  These
      are then output to a simple greyscale PBMPLUS PGM file.
   }
   \sstusage{
      ndf2pgm in out
   }
   \sstparameters{
      \sstsubsection{
         IN = NDF (Read)
      }{
         The name of the input NDF data structure (without the {\tt .sdf} 
         extension).  The suggested default is the current NDF if one exists,
         otherwise it is the current value.
      }
      \sstsubsection{
         OUT = \_CHAR (Read)
      }{
         The name of the PGM file be generated.
         The {\tt .pgm} name extension is added to any output filename that
         does not contain it.     
      }
   }
   \sstexamples{
      \sstexamplesubsection{
         ndf2pgm old new
      }{
         This converts the NDF called old (in file {\tt old.sdf}) to the
         PGM file {\tt new.pgm}.
      }
      \sstexamplesubsection{
         ndf2pgm in=spectre out=spectre.pgm
      }{
         This converts the NDF called spectre (in file {\tt spectre.sdf}) 
         to the PGM file {\tt spectre.pgm}.
      }
   }
   \sstnotes{
      This programme was written for diagnostic purposes and is included just
      in case someone finds it useful.
   }
   \sstimplementationstatus{
      Bad values in the data array are replaced with zero in the output
      PGM file.
   }
}

