\newpage
\sstroutine{
   TIFF2NDF
}{
   Converts a TIFF file into an NDF. 
}{
   \sstdescription{
      This Bourne shell script converts a 256 grey-level or b/w dithered
      Tag Image File Format (TIFF) into a Starlink NDF image format file.
      One or two-dimensional images can be handled.
      The script uses various PBMPLUS utilities to produce a FITS file 
      and then KAPPA applications FITSDIN, MATHS and FLIP to produce the
      final NDF. Error messages are converted into Starlink style (preceded
      by !).
   }
   \sstusage{
      tiff2ndf in [out]
   }
   \sstparameters{
      \sstsubsection{
         IN = FILENAME (Read)
      }{
         The name of the TIFF file to be converted (without the {\tt .tif}
         extension, which is assumed).
      }
      \sstsubsection{
         OUT = NDF (Write)
      }{
         The name of the NDF to be generated (without the {\tt .sdf} extension).
         If this is omitted, the value of the IN parameter is used.
      }
   }
   \sstexamples{
      \sstexamplesubsection{
         tiff2ndf old new
      }{
         This converts the TIFF file {\tt old.tif} into an NDF called new
         (in file {\tt new.sdf}).
      }
      \sstexamplesubsection{
         tiff2ndf horse
      }{
         This converts the TIFF file {\tt horse.tif} into an NDF called horse
         (in file {\tt horse.sdf}).
      }
   }
   \sstnotes{
      The following points should be remembered:
      \sstitemlist{
         \sstitem
            This initial version of the script handles only greyscale
            or b/w dithered images. Users are responsible for conversion
            of their images to this format prior to use. This ensures that
            the user is responsible for converting RGB values to brightness 
            values.
         \sstitem
            Input image file names must have the extension {\tt .tif}.
         \sstitem
            The PBMPLUS utilities {\tt tifftopnm}, {\tt ppmtopgm} and
            {\tt pgmtofits} must be available on your PATH.
      }
   }
   \sstdiytopic{
      Related Applications
   }{
      \CONVERT: NDF2TIFF.
   }
   \sstbugs{
      At the time of writing, this utility does not work on alpha OSF/1
      due to a problem with {\tt tifftopnm} in the standard release of 
      PBMPLUS. OSF/1 users should use a copy of {\tt tifftopnm} obtained from 
      Netpbm which is based on PBMPLUS but contains many improvements and
      additions. A suitable version may be obtained via WWW at
      \htmladdnormallink{{\tt http://www.astro.cf.ac.uk/pub/Grant.Privett/}}
      {http://www.astro.cf.ac.uk/pub/Grant.Privett/}.
      Alternatively, contact Grant Privett, {\tt gjp@astronomy.cardiff.ac.uk}.
   }
}

