\documentstyle{article} 
\pagestyle{myheadings}

%------------------------------------------------------------------------------
\newcommand{\stardoccategory}  {Starlink General Paper}
\newcommand{\stardocinitials}  {SGP}
\newcommand{\stardocnumber}    {25.9a}
\newcommand{\stardocauthors}   {M D Lawden}
\newcommand{\stardocdate}      {14 December 1990}
\newcommand{\stardoctitle}     {Starlink Site Manager's Guide --- Addendum}
%------------------------------------------------------------------------------

\newcommand{\stardocname}{\stardocinitials /\stardocnumber}
\renewcommand{\_}{{\tt\char'137}}     % re-centres the underscore
\markright{\stardocname}
\setlength{\textwidth}{160mm}
\setlength{\textheight}{240mm}
\setlength{\topmargin}{-5mm}
\setlength{\oddsidemargin}{0mm}
\setlength{\evensidemargin}{0mm}
\setlength{\parindent}{0mm}
\setlength{\parskip}{\medskipamount}
\setlength{\unitlength}{1mm}

\begin{document}
\thispagestyle{empty}
SCIENCE \& ENGINEERING RESEARCH COUNCIL \hfill \stardocname\\
RUTHERFORD APPLETON LABORATORY\\
{\large\bf Starlink Project\\}
{\large\bf \stardoccategory\ \stardocnumber}
\begin{flushright}
\stardocauthors\\
\stardocdate
\end{flushright}
\vspace{-4mm}
\rule{\textwidth}{0.5mm}
\vspace{5mm}
\begin{center}
{\Large\bf \stardoctitle}
\end{center}
\vspace{5mm}

\section{Data Protection Act}

Section 9 of SGP/25.9 stated that ``Starlink's files are believed to be
covered by the general SERC registration\ldots".
However, the DP officer at the University of Sussex was unsatisfied with
this assurance, and made further enquiries.
The result of this is that the Data Protection Registrar has advised the
University of Sussex that ``it, as a Starlink site, should ensure
that its registration covers use of Starlink as it is not covered by
SERC's registration".

The manager of every Starlink site should, therefore, inform the Data
Protection officer at that site of the existence and contents of any files
held on his Starlink computer(s) which contains information about specific
people.
Your DP officer will probably have a form for you to fill in.
File RLVAD::LADMINDIR:DPA.LIS contains the details that I submitted to our DP
officer at RAL with respect to our Starlink computers.
This may help you prepare your own submission.
You may also have local files that you should add to the list.
Users of Starlink facilities are also responsible for registering their
own files containing information on other people with their DP registrar.
However, I do not believe this involves the Site Manager.

\end{document}
