\documentstyle {article} 
\markright{SSN/29.3}
\setlength{\textwidth}{153mm}
\setlength{\textheight}{220mm}
\setlength{\oddsidemargin}{3mm}
\setlength{\evensidemargin}{3mm}
\pagestyle{myheadings}

\begin{document}
\thispagestyle{plain}
\noindent
SCIENCE \& ENGINEERING RESEARCH COUNCIL \hfill SSN/29.3\\
RUTHERFORD APPLETON LABORATORY\\
{\large\bf Starlink Project\\}
{\large\bf Starlink System Note 29.3}
\begin{flushright}
M D Lawden\\
3 June 1987
\end{flushright}
\vspace{-4mm}
\rule{\textwidth}{0.5mm}
\vspace{10mm}
\begin{center}
{\Large\bf SSL --- Startup, Shutdown, Login files}
\end{center}
\vspace{10mm}
\section {INTRODUCTION}
This note introduces the Startup, Shutdown and Login files used at Starlink
sites specifically to tailor the VMS system for the execution of Starlink
software.
\section {STARTUP FILES}
When a VAX/VMS system is booted, a sequence of command procedure files are
executed which prepare the software environment for users.
The full sequence is complicated and site-dependant, but as far as the Starlink
Software Collection is concerned the following components are the most
significant:
\setlength{\unitlength}{1mm}
\begin{center}
\begin{picture}(60,25)
\thicklines
\put (0,20){\framebox(60,5){SYS\$MANAGER:SYSTARTUP.COM}}
\put (0,10){\framebox(60,5){SSC:STARTUP.COM}}
\put (0,0){\framebox(60,5){LSSC:STARTUP.COM}}
\put (30,20){\vector(0,-1){5}}
\put (30,10){\vector(0,-1){5}}
\end{picture}
\end{center}
File SSC:STARTUP.COM defines some system logical names and installs Starlink
programs which require privilege.
This file is distributed with the Starlink Software Collection and should be
executed at every Starlink site.
Each site should also have a local startup command procedure in file
LSSC:STARTUP.COM which defines further system logical names required locally.
Each definition is identified with a particular software item name.
These files are subject to continual change so if you want to refer to the
latest version of them, print them out.
\section {SHUTDOWN FILE}
The command procedure in file SSC:SHUTDOWN.COM is executed when the system is
shut down.
At present, it does not do anything significant.
\section {LOGIN FILES}
When a user logs in to a Starlink VAX computer, the system attempts to execute a
command procedure specified in the user's authorization file.
This would normally be file LOGIN.COM in the users default directory.
If you wish to use Starlink software, include the following command in your
LOGIN.COM file:
\begin{verbatim}
         $ @SSC:LOGIN
\end{verbatim}
This command procedure defines some symbols required by users of the Starlink
Software Collection.
It also calls a local command procedure called LSSC:LOGIN.COM.
Thus, the call tree is:
\begin{center}
\begin{picture}(60,25)
\thicklines
\put (0,20){\framebox(60,5){[user default]LOGIN.COM}}
\put (0,10){\framebox(60,5){SSC:LOGIN.COM}}
\put (0,0){\framebox(60,5){LSSC:LOGIN.COM}}
\put (30,20){\vector(0,-1){5}}
\put (30,10){\vector(0,-1){5}}
\end{picture}
\end{center}
These command procedures are subject to continual change.
Their current contents can be examined at any time.
Each definition is identified with a particular software item name.
\end{document}
