\documentstyle{article}
\pagestyle{myheadings}

%------------------------------------------------------------------------------
\newcommand{\stardoccategory}  {Starlink System Note}
\newcommand{\stardocinitials}  {SSN}
\newcommand{\stardocnumber}    {27.5}
\newcommand{\stardocauthors}   {W F Lupton}
\newcommand{\stardocdate}      {9 January 1989}
\newcommand{\stardoctitle}     {HDS --- Hierarchical Data System}
%------------------------------------------------------------------------------

\newcommand{\stardocname}{\stardocinitials /\stardocnumber}
\markright{\stardocname}
\setlength{\textwidth}{160mm}
\setlength{\textheight}{240mm}
\setlength{\topmargin}{-5mm}
\setlength{\oddsidemargin}{0mm}
\setlength{\evensidemargin}{0mm}
\setlength{\parindent}{0mm}
\setlength{\parskip}{\medskipamount}
\setlength{\unitlength}{1mm}

\begin{document}
\thispagestyle{empty}
SCIENCE \& ENGINEERING RESEARCH COUNCIL \hfill \stardocname\\
RUTHERFORD APPLETON LABORATORY\\
{\large\bf Starlink Project\\}
{\large\bf \stardoccategory\ \stardocnumber}
\begin{flushright}
\stardocauthors\\
\stardocdate
\end{flushright}
\vspace{-4mm}
\rule{\textwidth}{0.5mm}
\vspace{5mm}
\begin{center}
{\Large\bf \stardoctitle}

\vspace {5mm}

{\large\bf Design and Implementation}
\end{center}
\vspace{5mm}

\setlength{\parskip}{0mm}
\tableofcontents
\listoftables
\listoffigures
\setlength{\parskip}{\medskipamount}
\markright{\stardocname}

\newpage
\section {INTENDED AUDIENCE}

This document is aimed at users of HDS who wish to know something about how it
is implemented, and at system programmers who are contemplating maintaining,
altering or adding new facilities to HDS. The reader is assumed to be familiar
with the VMS operating system, but no prior knowledge of HDS is assumed
(although the material will be easier to assimilate if he has used it as an
application programmer).

This is not an exhaustive document and, as is often the case, the ultimate
documentation is the code. However, all of the major concepts are discussed
here and anybody wishing seriously to get to grips with the internals of
HDS should definitely read it.

\section {HOW TO USE THIS DOCUMENT}

If you know nothing about HDS, read the INTRODUCTION and the HDS CONCEPTS
sections first. If you have already used it as an application programmer, you
can start at the HDS ARCHITECTURE section (if you know what you are looking for
you may be able to go straight there by referring to the CONTENTS). If you are
interested in modifying HDS then you can start at the HDS IMPLEMENTATION
section. If you just want some tips for making more efficient use of HDS then
you can start at the TUNING section.

Throughout the text, footnotes are used to draw attention to areas where
the implementation requires further work or where something is handled in
an unsatisfactory way. Footnotes are not used for any other purpose.

\section {INTRODUCTION}

HDS is a hierarchical data storage and retrieval system. It is structured as a
set of subroutines which are used by application programs to create, store,
access and manipulate {\em data objects}. These objects have a tree-like
structure which can be altered in shape as well as content during program
execution. It is this dynamic structuring facility which most clearly
differentiates HDS from normal file systems. The main features of HDS are:

\begin {itemize}
\item Self description of objects
\item Hierarchical structure of objects
\item Dynamic alteration of object shape
\item Automatic type conversion
\item Bad value identification
\item Mapping and caching to improve efficiency
\end {itemize}

HDS uses no system-wide resources such as global sections or event flags and
does not require any VMS process privileges to execute. With few exceptions,
the guidelines recommended in \cite{creating_modular_procedures} have been
adhered to. It follows the VAX Procedure Calling Standard and can thus be used
by any DEC-supported language product. VAX FORTRAN is the programming language
normally used to access the system. It is described in
\cite{programming_in_fortran} and the routine specifications are given in
\cite{sun_92}.

\section {HDS CONCEPTS}

The basic building block of data stored by the system is the {\em data object}
which will be referred to by the name {\em object}. An object possesses the
following attributes:

\begin {description}
\item [NAME]: a character string which identifies the object. Although in
general this is arbitrary, special meaning {\em may} be attached to certain
names by application software or environment.
\item [TYPE]: a character string which specifies the object's {\em type}.
There are two {\em classes}:
\begin {itemize}
\item {\bf Primitive} --- actual data values such as a 2-D image array, a
character string denoting the name of a telescope, or a pair of integer
coordinate values.
\item {\bf Structure} --- collections of other objects (primitive or structure)
which are grouped together because of an inherent logical relationship.
\end {itemize}
\item [SHAPE]: an n-element array specifying the sizes of the n dimensions of
the object. The maximum number of dimensions is seven. A scalar object is
regarded as having zero dimensions.
\item [STATE]: a logical variable which specifies whether or not data values
are defined for the object.
\end {description}

Structure objects are tree-like and each object in them can be a scalar or an
array with up to 7 dimensions. Figure~\ref{diagrams_of_structure_objects}
contains some diagrams of structure objects.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(130,43)
\thicklines
\put (7,40){P(n)}
\put (67,40){S(n)}
\put (67,30){P(n)}
\put (82,30){P(n)}
\put (97,30){P(n)}
\put (7,30){S(n)}
\put (7,20){P(n)}
\put (22,20){S(n)}
\put (37,20){P(n,n,n)}
\put (22,10){P(n)}
\put (37,10){P(n)}
\put (52,10){S(n,n)}
\put (52,0){P(n)}
\put (67,0){P(n,n,n,n)}
\put (70,38){\line(0,-1){4}}
\put (70,36){\line(1,0){30}}
\put (85,36){\line(0,-1){2}}
\put (100,36){\line(0,-1){2}}
\put (10,28){\line(0,-1){4}}
\put (10,26){\line(1,0){30}}
\put (25,26){\line(0,-1){2}}
\put (40,26){\line(0,-1){2}}
\put (25,18){\line(0,-1){4}}
\put (25,16){\line(1,0){30}}
\put (40,16){\line(0,-1){2}}
\put (55,16){\line(0,-1){2}}
\put (55,8){\line(0,-1){4}}
\put (55,6){\line(1,0){15}}
\put (70,6){\line(0,-1){2}}
\put (100,15){\bf KEY}
\put (90,10){S(n) --- Structure object}
\put (90,5){P(n) --- Primitive object}
\end {picture}
\caption {Diagrams of Structure Objects}
\label {diagrams_of_structure_objects}
\end {center}
\end {figure}

\subsection {Primitive Objects}

Primitive objects contain the actual data values that are processed by
application software. HDS supports five {\em standard} primitive data types
which correspond to the main FORTRAN-77 types. They are illustrated in
Table~\ref{standard_primitive_data_types}.

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|l|l||}
\hline
Primitive data type	& VAX FORTRAN type		& HDS type \\
\hline
Integer			& {\tt INTEGER}			& {\tt `\_INTEGER'} \\
Single floating point	& {\tt REAL}			& {\tt `\_REAL'} \\
Double floating point	& {\tt DOUBLE PRECISION}	& {\tt `\_DOUBLE'} \\
Logical			& {\tt LOGICAL}			& {\tt `\_LOGICAL'} \\
Character		& {\tt CHARACTER[*n]}		& {\tt `\_CHAR[*n]'} \\
\hline
\end {tabular}
\caption {Standard Primitive Data Types}
\label {standard_primitive_data_types}
\end {center}
\end {table}

Because of the diverse nature of astronomical data a further set of primitive
data types called {\em non-standard} are also supported but their
indiscriminate use may make software less portable. They are illustrated
in Table~\ref{non_standard_primitive_data_types}.

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|l|l||}
\hline
Primitive data type	& VAX FORTRAN type	& HDS type \\
\hline
Byte			& {\tt BYTE}		& {\tt `\_BYTE'} \\
Unsigned byte		& {\tt BYTE}		& {\tt `\_UBYTE'} \\
Word			& {\tt INTEGER*2}	& {\tt `\_WORD'} \\
Unsigned word		& {\tt INTEGER*2}	& {\tt `\_UWORD'} \\
\hline
\end {tabular}
\caption {Non-standard Primitive Data Types}
\label {non_standard_primitive_data_types}
\end {center}
\end {table}

Note the use of the leading underscore in the HDS type names --- this is used
to distinguish between primitive and structure objects. The precision of
primitive data depends on the architecture of the host machine. The only host
machines used at present are those with the DEC VAX architecture. The VAX data
types are described in \cite{programming_in_fortran} and their sizes and ranges
are illustrated in Table~\ref{sizes_and_ranges_of_primitive_data_types}.

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|r|l||}
\hline
Data type	& Storage	& Range \\
\hline
{\tt INTEGER}	& 32 bits	& {\tt -2147483648 to 2147483647} \\
{\tt REAL}	& 32 bits	& {\tt +/- 0.29e-38 to 1.7e+38} \\
{\tt DOUBLE}	& 64 bits	& {\tt +/- 0.29d-38 to 1.7d+38} \\
{\tt LOGICAL}	& 32 bits	& {\tt TRUE or FALSE} \\
{\tt CHAR}	& 8 bits	& {\tt Any 8-bit pattern} \\
{\tt CHAR*n}	& n*8 bits	& {\tt 0 < n < 65536} \\
\hline
{\tt BYTE}	& 8 bits	& {\tt -128 to 127} \\
{\tt UBYTE}	& 8 bits	& {\tt 0 to 255} \\
{\tt WORD}	& 16 bits	& {\tt -32768 to 32767} \\
{\tt UWORD}	& 16 bits	& {\tt 0 to 65535} \\
\hline
\end {tabular}
\caption {Sizes and Ranges of Primitive Data Types}
\label {sizes_and_ranges_of_primitive_data_types}
\end {center}
\end {table}

The REAL and DOUBLE data types correspond to the VAX F\_floating and
D\_floating formats, which have a precision of 7 and 16 decimal places
respectively; the extended range G\_floating and H\_floating formats are
not currently supported.

\subsubsection {Type Conversion}

Generally an application program need not concern itself with the actual data
type of the primitive objects it processes since HDS performs automatic data
conversion when necessary. Although this is transparent to a program, it may be
rather more obvious to a user sitting at a terminal, particularly when large
arrays are being processed. In many circumstances it may be more efficient to
provide special-case code for each data type to avoid the overheads involved
with automatic conversion. An alternative is to encourage users to bias their
data towards one specific format such as single precision floating point;
application software can then be designed around this chosen type.

Data of one numeric type can be converted to any of the other numeric types
although precision may be lost in some cases. The conversion is performed using
ordinary C assignment statements. When arithmetic overflow occurs the data item
becomes undefined and is replaced by a {\em bad} value. These are illustrated
in Table~\ref{bad_values_for_primitive_data_types}\footnote{The data conversion
routines are currently unsatisfactory: a bad value will not be converted to a
bad value, overflows on DOUBLE PRECISION and REAL to INTEGER and LOGICAL are
not detected, and they are relatively slow.}.

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|l|l||}
\hline
Data type		& Bad value		& Hexadecimal pattern \\
\hline
{\tt INTEGER}		& {\tt -2147483648}	& {\tt `80000000'} \\
{\tt REAL}		& {\tt -1.7014117e+38}	& {\tt `FFFFFFFF'} \\
{\tt DOUBLE PRECISION}	& {\tt -1.7014117e+38}	& {\tt `FFFFFFFFFFFFFFFF'} \\
{\tt LOGICAL}		& {\tt integer -1}	& {\tt `FFFFFFFF'} \\
{\tt CHAR[*n]}		& {\tt `*...*'}		& {\tt `2A...2A'} \\
{\tt BYTE}		& {\tt -128}		& {\tt `80'} \\
{\tt UBYTE}		& {\tt 255}		& {\tt `FF'} \\
{\tt WORD}		& {\tt -32768}		& {\tt `8000'} \\
{\tt UWORD}		& {\tt 65535}		& {\tt `FFFF'} \\
\hline
\end {tabular}
\caption {Bad Values for Primitive Data Types}
\label {bad_values_for_primitive_data_types}
\end {center}
\end {table}

The conversion process is {\em not} prematurely terminated when overflow is
detected --- the condition is ultimately signalled through the HDS subroutine
status return.

The two non-numeric types LOGICAL and CHARACTER are converted to and from each
other and the numeric types using the following rules:

\begin {itemize}
\item LOGICAL TRUE and FALSE are converted to character strings
{\tt `TRUE'} and {\tt `FALSE'} and numbers 1 and 0 respectively.
\item Character strings beginning with any of the characters {\tt \{t,T,y,Y\}}
are converted to LOGICAL TRUE. All other character strings are converted to
LOGICAL FALSE.
\item Numbers are converted to LOGICALs by first converting to an integer
and then looking only at the least significant bit. If it is set, the value
is TRUE and if not it is FALSE.
\item Numbers are converted to and from character strings using the {\tt
LIB\$CVT\_DX\_DX} routine --- see \cite{run_time_library_reference}.
\end {itemize}

Truncation or padding may occur between character-type objects of different
lengths. When destination strings are longer than source strings the values
are padded to the right with spaces. However, truncation can result in
significant characters being lost, which is regarded as an error by HDS (unlike
FORTRAN-77). For example, consider the following 4-element {\tt \_CHAR*47}
object (courtesy of Lewis Carroll):

\begin {verbatim}
        `Yet what are all such gaieties to me           '
        `Whose thoughts are full of indices and surds?  '
        `x-squared plus seven-x plus fifty-three        '
        `Equals eleven thirds                           '
\end{verbatim}

An HDS program attempting to read this into a {\tt \_CHAR*38} array would find
that although the first and last strings are transferred successfully (because
only spaces were discarded), an error status would be returned because
significant characters were lost from the second and third strings.
Character-to-character translation is never prematurely terminated; this allows
the program to ignore the truncation errors if this is sensible.

\subsection {Structure Objects}

In a typical astronomical data-reduction environment it is unlikely that
primitive objects will exist as independent entities. Consider a plate-scan of
the galaxy NGC1365 which has been reduced to a 2-D array of floating point
values. The supplementary information accompanying the image might be:

\begin {verbatim}
        TELESCOPE  = `UKST'
        INSTRUMENT = `3A0/PDS'
        RA         = 10 59 17.107
        DEC        = +47 23 15.94
        EQUINOX    = B1950.0
\end{verbatim}

Image-processing systems have tended to store such information in rigid fixed
formats. However, HDS gives you the flexibility to structure it in whatever
format you regard as natural. Figure~\ref{an_example_hierarchical_structure}
illustrates one possible arrangement. The top-level object is called NGC1365
and is of type {\em plate} (type names will be expressed in lowercase and
enclosed in brackets). The structure is held in a single VMS file called a {\em
container file}. Every container file has a top-level object which would
usually (but not necessarily) have the same name as the VMS file. Since
structures are built top-down, if a file is initially created with a primitive
top-level object no more objects can subsequently be added.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(150,110)
\thicklines
\put (60,105){\framebox(26,5){NGC1365}}
\put (60,100){\framebox(26,5){[plate]}}
\put (0,85){\framebox(26,5){SCAN\_DATA}}
\put (0,80){\framebox(26,5){[\_real]}}
\put (0,75){\framebox(26,5){(512,512)}}
\put (40,85){\framebox(26,5){TELESCOPE}}
\put (40,80){\framebox(26,5){[\_char*4]}}
\put (40,75){\framebox(26,5){`UKST'}}
\put (80,85){\framebox(26,5){INSTRUMENT}}
\put (80,80){\framebox(26,5){[\_char*7]}}
\put (80,75){\framebox(26,5){`3AO/PDS'}}
\put (120,85){\framebox(26,5){SKY\_POSITION}}
\put (120,80){\framebox(26,5){[equatorial]}}
\put (30,60){\framebox(26,5){RA}}
\put (30,55){\framebox(26,5){[hms]}}
\put (80,60){\framebox(26,5){DEC}}
\put (80,55){\framebox(26,5){[dms]}}
\put (120,60){\framebox(26,5){EQUINOX}}
\put (120,55){\framebox(26,5){[equinox]}}
\put (0,40){\framebox(26,5){HOURS}}
\put (0,35){\framebox(26,5){[\_byte]}}
\put (0,30){\framebox(26,5){10}}
\put (30,40){\framebox(26,5){MINUTES}}
\put (30,35){\framebox(26,5){[\_byte]}}
\put (30,30){\framebox(26,5){59}}
\put (60,40){\framebox(26,5){SECONDS}}
\put (60,35){\framebox(26,5){[\_real]}}
\put (60,30){\framebox(26,5){17.107}}
\put (100,40){\framebox(26,5){SYSTEM}}
\put (100,35){\framebox(26,5){[\_char*1]}}
\put (100,30){\framebox(26,5){`B'}}
\put (130,40){\framebox(26,5){EPOCH}}
\put (130,35){\framebox(26,5){[\_real]}}
\put (130,30){\framebox(26,5){1950.0}}
\put (50,15){\framebox(26,5){DEGREES}}
\put (50,10){\framebox(26,5){[\_byte]}}
\put (50,5){\framebox(26,5){47}}
\put (80,15){\framebox(26,5){ARCMINS}}
\put (80,10){\framebox(26,5){[\_byte]}}
\put (80,5){\framebox(26,5){23}}
\put (110,15){\framebox(26,5){ARCSECS}}
\put (110,10){\framebox(26,5){[\_real]}}
\put (110,5){\framebox(26,5){15.94}}
\put (13,95){\line(1,0){120}}
\put (43,70){\line(1,0){90}}
\put (13,50){\line(1,0){60}}
\put (113,50){\line(1,0){30}}
\put (63,25){\line(1,0){60}}
\put (73,100){\line(0,-1){5}}
\put (133,55){\line(0,-1){5}}
\put (13,95){\vector(0,-1){5}}
\put (53,95){\vector(0,-1){5}}
\put (93,95){\vector(0,-1){5}}
\put (133,95){\vector(0,-1){5}}
\put (43,70){\vector(0,-1){5}}
\put (93,70){\vector(0,-1){5}}
\put (133,80){\vector(0,-1){15}}
\put (13,50){\vector(0,-1){5}}
\put (43,55){\vector(0,-1){10}}
\put (73,50){\vector(0,-1){5}}
\put (93,55){\vector(0,-1){35}}
\put (113,50){\vector(0,-1){5}}
\put (143,50){\vector(0,-1){5}}
\put (63,25){\vector(0,-1){5}}
\put (123,25){\vector(0,-1){5}}
\end {picture}
\caption {An Example Hierarchical Structure}
\label {an_example_hierarchical_structure}
\end {center}
\end {figure}

In Figure~\ref{an_example_hierarchical_structure}, NGC1365 is a structure with
four {\em components}, three of which are primitive objects and the fourth a
structure. RA, DEC and EQUINOX are also structures which store the values as
discrete numeric items. Rather than separating their individual parts you could
have chosen to store them as primitive character objects as illustrated in
Figure~\ref{an_alternative_representation_for_sky_position}, although this
scheme would obviously require software within the application program to decode
the strings if the individual items were to be used arithmetically.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(110,35)
\thicklines
\put (40,30){\framebox(30,5){SKY\_POSITION}}
\put (40,25){\framebox(30,5){[equatorial]}}
\put (0,10){\framebox(30,5){RA}}
\put (0,5){\framebox(30,5){[\_char*12]}}
\put (0,0){\framebox(30,5){`10:59:17.107'}}
\put (40,10){\framebox(30,5){DEC}}
\put (40,5){\framebox(30,5){[\_char*12]}}
\put (40,0){\framebox(30,5){`47:23:15.94'}}
\put (80,10){\framebox(30,5){EQUINOX}}
\put (80,5){\framebox(30,5){[\_char*7]}}
\put (80,0){\framebox(30,5){`B1950.0'}}
\put (15,20){\line(1,0){80}}
\put (15,20){\vector(0,-1){5}}
\put (55,25){\vector(0,-1){10}}
\put (95,20){\vector(0,-1){5}}
\end {picture}
\caption {An Alternative Representation for SKY\_POSITION}
\label {an_alternative_representation_for_sky_position}
\end {center}
\end {figure}

In reality, conventions will be adopted to promote standardisation and
portability. For example, any character-type primitive object named RA might be
expected to contain a right ascension value of the form {\em
hours:minutes:seconds}. Alternatively, any object of type {\em hms} may be
assumed to be a structure with three primitive components named HOURS, MINUTES
and SECONDS.

Applications should not attach any significance to the order in which
components appear in a structure. Currently, new components are simply appended
to the existing list of objects when they are created although this is an
implementation strategy not a design feature\footnote{The algorithms used
for creating, deleting and searching for objects are relatively
unsophisticated. All searches are simply linear.}.

So far the examples have illustrated the use of scalar structures. However, a
structure can also have a {\em shape}, ie.\ have one or more dimensions. A
typical application of structure arrays is the processing of a sequential set
of data files held on magnetic tape. For example,
Figure~\ref{an_example_of_a_structure_array} shows how a set of raw IPCS images
could be stored as a 1-D structure. Given a tape labelled AAO28, a program
could create a structure array in a container file and subsequently build up
the component objects in each structure element as the tape is processed.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(120,25)
\thicklines
\put (0,20){\framebox(30,5){AAO28(1)}}
\put (0,15){\framebox(30,5){[ipcs\_raw]}}
\put (30,20){\framebox(30,5){AAO28(2)}}
\put (30,15){\framebox(30,5){[ipcs\_raw]}}
\put (60,20){\framebox(30,5){AAO28(3)}}
\put (60,15){\framebox(30,5){[ipcs\_raw]}}
\put (90,20){\framebox(30,5){AAO28(4)}}
\put (90,15){\framebox(30,5){[ipcs\_raw]}}
\put (5,3){\framebox(8,2){\ldots}}
\put (17,3){\framebox(8,2){\ldots}}
\put (35,3){\framebox(8,2){\ldots}}
\put (47,3){\framebox(8,2){\ldots}}
\put (65,3){\framebox(8,2){\ldots}}
\put (77,3){\framebox(8,2){\ldots}}
\put (95,3){\framebox(8,2){\ldots}}
\put (107,3){\framebox(8,2){\ldots}}
\put (9,10){\line(1,0){12}}
\put (39,10){\line(1,0){12}}
\put (69,10){\line(1,0){12}}
\put (99,10){\line(1,0){12}}
\put (15,15){\line(0,-1){5}}
\put (45,15){\line(0,-1){5}}
\put (75,15){\line(0,-1){5}}
\put (105,15){\line(0,-1){5}}
\put (9,10){\vector(0,-1){5}}
\put (21,10){\vector(0,-1){5}}
\put (39,10){\vector(0,-1){5}}
\put (51,10){\vector(0,-1){5}}
\put (69,10){\vector(0,-1){5}}
\put (81,10){\vector(0,-1){5}}
\put (99,10){\vector(0,-1){5}}
\put (111,10){\vector(0,-1){5}}
\end {picture}
\caption {An Example of a Structure Array}
\label {an_example_of_a_structure_array}
\end {center}
\end {figure}

As with the naming and typing of objects, conventions will also be adopted for
the shaping of objects, eg.\ whether all tape-reading utilities should create
structure arrays as above or a (non-related) set of scalar structures.

\subsection {Temporary Objects}

HDS allows you to create temporary objects. A temporary object is exactly the
same as a permanent object --- the only difference is that temporary objects
are created in a container file that is automatically deleted when it is
closed. This is all handled automatically for you. When the first temporary
object is created, a file called {\tt SYS\$SCRATCH:Tnnnnnnnn.SDF}\footnote{Note
that this is often the default login directory.} is created ({\tt nnnnnnnn} is
the HEX VMS process ID which is used to prevent any naming conflict with the
user's permanent files). The first temporary object is called {\tt
TEMP\_00000000}, the second is called {\tt TEMP\_00000001} and so on.

\subsection {Locators}
\label {locators}

The application routines use entities called {\em locator}\/s to refer to HDS
objects. From the point of view of an application program, a locator is a {\tt
CHARACTER*15} variable. From the point of view of HDS it looks like the data
structure illustrated in Figure~\ref{structure_of_a_locator}.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,25)
\thicklines
\put (0,0){\framebox(40,5){reserve}}
\put (0,5){\framebox(40,5){pointer to LCP}}
\put (0,10){\framebox(40,5){validity check}}
\put (42,2){(7 bytes)}
\put (42,7){(4 bytes)}
\put (42,12){(4 bytes)}
\end {picture}
\caption {Structure of a Locator}
\label {structure_of_a_locator}
\end {center}
\end {figure}

An active locator contains a pointer to an associated {\em Locator Control
Packet} (LCP), which resides on a doubly-linked list maintained by HDS. When a
new locator is required, a packet is dynamically allocated from process space
and added to the list. HDS actually maintains a look-aside list of free LCPs
in order to reduce the overheads involved in allocating dynamic memory. An LCP
acts as a cache for information about the object that would otherwise have to
be fetched from disc. This includes the object's name, type and shape, fields
which describe the attributes of any primitive data that is currently mapped,
and various flags which indicate the state of the object.

Further discussion of the use of the LCP is deferred until
Section~\ref{use_of_locators} by which time the appropriate terminology will
have been introduced.

\section {HDS ARCHITECTURE}

HDS is implemented as three distinct modules --- their relationship is
illustrated in Figure~\ref{relationship_of_hds_modules}. The DAT module
contains the user-callable routines. The REC module contains the record
management routines. The EXC module contains the exception handlers (which are
also user-callable).

\begin {figure}[htbp]
\begin {center}
\begin {picture}(50,20)
\thicklines
\put (0,15){\framebox(20,5){DAT}}
\put (0,0){\framebox(20,5){REC}}
\put (30,0){\framebox(20,5){EXC}}
\put (10,15){\vector(0,-1){10}}
\put (10,15){\vector(3,-1){29}}
\put (20,2.5){\vector(1,0){10}}
\end {picture}
\caption {Relationship of HDS Modules}
\label {relationship_of_hds_modules}
\end {center}
\end {figure}

The modules have been made independent of each other to facilitate further
development of the system. For example, the REC routines do not recognise HDS
objects --- they are called by the DAT module to manipulate records within a
container file. DAT and REC are two different layers between the application
program and the underlying file system. Each provides the layer above with a
characteristic data entity to manipulate. This is illustrated in
Figure~\ref{entities_accessed_at_different_levels}.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,50)
\thicklines
\put (0,0){\framebox(40,5){VAX/VMS}}
\put (0,15){\framebox(40,5){REC}}
\put (0,30){\framebox(40,5){DAT}}
\put (0,45){\framebox(40,5){Application program}}
\put (20,10){\makebox(0,0){Container file, Block}}
\put (20,25){\makebox(0,0){Record}}
\put (20,40){\makebox(0,0){HDS Data Object}}
\put (20,38){\vector(0,-1){3}}
\put (20,23){\vector(0,-1){3}}
\put (20,8){\vector(0,-1){3}}
\put (20,45){\line(0,-1){3}}
\put (20,30){\line(0,-1){3}}
\put (20,15){\line(0,-1){3}}
\put (50,35){\line(0,-1){20}}
\put (50,35){\line(-1,0){2}}
\put (50,15){\line(-1,0){2}}
\put (50,25){\line(1,0){2}}
\put (56,25){\makebox(0,0){HDS}}
\end {picture}
\caption {Entities Accessed at Different Levels}
\label {entities_accessed_at_different_levels}
\end {center}
\end {figure}

\subsection {Objects and Records}

The DAT module currently defines four classes of records:

\begin {description}
\item [CONTAINER record]: points to the file's top-level object.
\item [COMPONENT record]: lists a structure's components.
\item [STRUCTURE record]: describes a structure-type object.
\item [PRIMITIVE record]: describes a primitive-type object.
\end {description}

The REC module does not distinguish between different record classes. New ones
can be defined in DAT without having to change REC.

Figure~\ref{how_an_object_maps_onto_records} shows an example of how an object
structure is mapped onto a record structure by HDS.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(140,72)
\thicklines
\put (0,69){{\bf Object Structure}}
\put (0,29){{\bf Record Structure}}
\put (20,55){S: Structure}
\put (20,50){P: Primitive}
\put (20,15){Co: Container}
\put (20,10){C: Component}
\put (20,5){S: Structure}
\put (20,0){P: Primitive}
\put (48,69){S(n)}
\put (78,69){S(n)}
\put (78,59){P(n)}
\put (108,69){P(n)}
\put (108,59){P(n)}
\put (108,49){S(n)}
\put (138,49){P(n)}
\put (138,39){P(n)}
\put (38,29){Co}
\put (48,29){S(n)}
\put (68,29){C}
\put (78,29){S(n)}
\put (98,29){C}
\put (108,29){P(n)}
\put (78,19){P(n)}
\put (108,19){P(n)}
\put (108,9){S(n)}
\put (128,9){C}
\put (138,9){P(n)}
\put (138,-1){P(n)}
\put (55,70){\line(1,0){22}}
\put (85,70){\line(1,0){22}}
\put (75,60){\line(1,0){2}}
\put (105,60){\line(1,0){2}}
\put (105,50){\line(1,0){2}}
\put (115,50){\line(1,0){22}}
\put (135,40){\line(1,0){2}}
\put (43,30){\line(1,0){5}}
\put (55,30){\line(1,0){12}}
\put (72,30){\line(1,0){5}}
\put (85,30){\line(1,0){12}}
\put (102,30){\line(1,0){5}}
\put (75,20){\line(1,0){2}}
\put (105,20){\line(1,0){2}}
\put (105,10){\line(1,0){2}}
\put (115,10){\line(1,0){12}}
\put (132,10){\line(1,0){5}}
\put (135,0){\line(1,0){2}}
\put (75,70){\line(0,-1){10}}
\put (105,70){\line(0,-1){20}}
\put (135,50){\line(0,-1){10}}
\put (75,30){\line(0,-1){10}}
\put (105,30){\line(0,-1){20}}
\put (135,10){\line(0,-1){10}}
\end {picture}
\caption {How an Object Maps on to Records}
\label {how_an_object_maps_onto_records}
\end {center}
\end {figure}

Elements which are arrays are indicated by appending {\em (n)} to the symbol;
in general, these can have up to seven dimensions. Each array element has the
structure represented by the tree to its right. Note that {\em each} element of
a structure array in the {\bf Object Structure} diagram corresponds in the {\bf
Record Structure} diagram to a Component record which points to Structure
and/or Primitive records that represent the component objects. This means
that elements of a structure array do not automatically have all the same
components as each other.

\subsection {Files and Blocks}

Container files are organised as a sequence of 512-byte fixed length blocks
which correspond directly to disc blocks and VAX/VMS pages. Blocks are the
basic unit of disc space allocation. There are three types of block:

\begin {description}
\item [HCB]: Header Control Block
\item [LRB]: Logical Record Block
\item [PDB]: Pure Data Block
\end {description}

Figure~\ref{structure_of_a_container_file} illustrates how they are arranged in
a container file.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,25)
\thicklines
\put (0,0){\framebox(40,5){LRB or PDB}}
\put (0,10){\framebox(40,5){LRB or PDB}}
\put (0,15){\framebox(40,5){LRB}}
\put (0,20){\framebox(40,5){HCB}}
\put (0,10){\line(0,-1){1}}
\put (0,8){\line(0,-1){1}}
\put (0,6){\line(0,-1){1}}
\put (40,10){\line(0,-1){1}}
\put (40,8){\line(0,-1){1}}
\put (40,6){\line(0,-1){1}}
\end {picture}
\caption {Structure of a Container File}
\label {structure_of_a_container_file}
\end {center}
\end {figure}

The {\em Header Control Block} is always the first block in every container
file. It contains the following fields and is illustrated in
Figure~\ref{structure_of_the_header_control_block}.

\begin {itemize}
\item {\bf File stamp} --- an arbitrary bit-string used by the REC module to
identify the file as a valid HDS container file.
\item {\bf Version number} --- used as a check against the current version of
HDS when an existing container file is opened. An error is signalled if the
file version number exceeds the software version number (which is only ever
incremented when the file format is changed).
\item {\bf End-of-file block number} --- the last physical block in the
file; it indicates the total file space.
\item {\bf Free space stack} --- contains a finite number of entries which
specify how much free space there is in Logical Record Blocks and which Pure
Data Blocks are free. This is discussed in Section~\ref{disc_space_management}.
\end {itemize}

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,25)
\thicklines
\put (0,0){\framebox(40,5){free space stack}}
\put (0,5){\framebox(40,5){reserve}}
\put (0,10){\framebox(40,5){end-of-file block no.}}
\put (0,15){\framebox(40,5){HDS version no.}}
\put (0,20){\framebox(40,5){file-stamp}}
\put (42,2){(480 bytes)}
\put (42,7){(24 bytes)}
\put (42,12){(32 bits)}
\put (42,17){(8 bits)}
\put (42,22){(24 bits)}
\end {picture}
\caption {Structure of the Header Control Block}
\label {structure_of_the_header_control_block}
\end {center}
\end {figure}

A {\em Logical Record Block} is partitioned into a number of 34-byte {\em
chips}. A chip is the unit used by the REC module to allocate space in an LRB
for a record. Each record occupies an integral number of chips, the number
depending on its size (blocks are not spanned). With 15 chips per block, this
leaves 2 bytes for a chip bitmap which specifies which chips are currently
allocated (bit-set) and which are free (bit-clear). An LRB is illustrated
in Figure~\ref{structure_of_a_logical_record_block}.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,25)
\thicklines
\put (0,0){\framebox(40,5){chip [14]}}
\put (0,10){\framebox(40,5){chip [1]}}
\put (0,15){\framebox(40,5){chip [0]}}
\put (0,20){\framebox(40,5){bitmap}}
\put (0,10){\line(0,-1){1}}
\put (0,8){\line(0,-1){1}}
\put (0,6){\line(0,-1){1}}
\put (40,10){\line(0,-1){1}}
\put (40,8){\line(0,-1){1}}
\put (40,6){\line(0,-1){1}}
\put (42,22){(2 bytes)}
\put (42,17){(34 bytes)}
\put (42,12){(34 bytes)}
\put (42,2){(34 bytes)}
\end {picture}
\caption {Structure of a Logical Record Block}
\label {structure_of_a_logical_record_block}
\end {center}
\end {figure}

A {\em Pure Data Block} is used to store data and, from the point of view of
HDS, can simply be regarded as a sequence of 512 bytes. PDB space is allocated
as a {\em frame} which is a contiguous sequence of PDBs.

\subsection {Record Structures}

Each record within a file can be located by a unique {\em Record-ID} (RID)
which is unchanging for the lifetime of the record. This is the relative block
number within the file and the chip offset within the block. Allowing 20 bits
for the block number gives a maximum file space of 1048576 blocks (512 Mbytes)
which should not prove too restrictive. The structure of a RID is illustrated
in Figure~\ref{structure_of_a_record_id}.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,10)
\thicklines
\put (0,5){\framebox(40,5){block no.}}
\put (0,0){\framebox(40,5){chip offset}}
\put (42,7){(20 bits)}
\put (42,2){(4 bits)}
\end {picture}
\caption {Structure of a Record-ID}
\label {structure_of_a_record_id}
\end {center}
\end {figure}

Every record has a unique parent (owner) record which is also identified by its
RID. However, unlike a record's personal identifier which it keeps for life,
its parent-ID can be modified which gives the effect of moving the record
(and all records owned by it) from one location in a structure to another.

A RID occupies 24 bits. Occasionally one sees a RID apparently occupying
32 bits. This is because the DAT module knows nothing about the structure
of a RID and always regards it as a 32-bit integer.

A Record consists of three {\em domains}, illustrated in
Figure~\ref{structure_of_a_record}:

\begin {description}
\item [CONTROL domain]: holds the Record Control Label (RCL).
This contains the record size and class, the domain lengths, the parent
Record-ID, and various bit flags describing record attributes.
\item [STATIC domain]: holds small amounts of data that are static in length
(although not necessarily in content).
\item [DYNAMIC domain]: holds large amounts of data or data whose length is
dynamic in nature.
\end {description}

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,15)
\thicklines
\put (0,10){\framebox(40,5){Control domain (RCL)}}
\put (0,5){\framebox(40,5){Static domain}}
\put (0,0){\framebox(40,5){Dynamic domain}}
\put (42,12){(10 bytes)}
\put (42,7){($\leq $498 bytes)}
\put (42,2){($\leq 2^{32}$ bytes)}
\end {picture}
\caption {Structure of a Record}
\label {structure_of_a_record}
\end {center}
\end {figure}

The {\em Record Control Label} (RCL) is illustrated in
Figure~\ref{structure_of_a_record_control_label}. The 4 flags A, C, M and Z
signify the following when set:

\begin {description}
\item [A]: domain (D) is active, ie.\ the contents are meaningful.
\item [C]: domain (D) is chained --- see Section~\ref{chained_dynamic_domains}.
\item [M]: domain (D) is currently mapped for write-access.
\item [Z]: domain (D) is to be explicitly zeroed when created.
\end {description}

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,25)
\thicklines
\put (0,20){\framebox(40,5){parent RID}}
\put (0,15){\framebox(25,5){reserve}}
\put (25,15){\framebox(5,5){M}}
\put (30,15){\framebox(5,5){Z}}
\put (35,15){\framebox(5,5){A}}
\put (0,10){\framebox(5,5){C}}
\put (5,10){\framebox(15,5){class}}
\put (20,10){\framebox(20,5){size}}
\put (0,5){\framebox(40,5){domain (S) length}}
\put (0,0){\framebox(40,5){domain (D) length}}
\put (42,22){(24 bits)}
\put (42,17){($5+1+1+1$ bits)}
\put (42,12){($1+3+4$ bits)}
\put (42,7){(8 bits)}
\put (42,2){(32 bits)}
\end {picture}
\caption {Structure of a Record Control Label}
\label {structure_of_a_record_control_label}
\end {center}
\end {figure}

\subsubsection {Chained Dynamic Domains}
\label {chained_dynamic_domains}

If the initial size of the record prevents it from being stored in an LRB, the
dynamic domain is {\em chained} --- a frame of PDBs is allocated from the free
space stack to hold the actual data and the position of the first block is
stored in the record's dynamic domain. An existing unchained record can become
chained due to the subsequent expansion of the dynamic domain. To illustrate
the basic characteristics of chained and unchained records, consider the
configurations illustrated in
Figure~\ref{an_example_of_a_chained_and_an_unchained_record}.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(100,65)
\thicklines
\put (0,55){\framebox(40,5){RCL}}
\put (0,50){\framebox(40,5){domain S}}
\put (0,45){\framebox(40,5){*}}
\put (0,30){\framebox(40,5){\ldots}}
\put (0,25){\framebox(40,5){\ldots}}
\put (0,20){\framebox(40,5){\ldots}}
\put (0,15){\framebox(40,5){\ldots}}
\put (0,10){\framebox(40,5){\ldots}}
\put (0,5){\framebox(40,5){\ldots}}
\put (0,0){\framebox(40,5){\ldots}}
\put (60,55){\framebox(40,5){RCL}}
\put (60,50){\framebox(40,5){domain S}}
\put (60,45){\framebox(40,5){domain D}}
\put (20,47.5){\vector(0,-1){12.5}}
\put (42,0){\vector(0,1){35}}
\put (42,35){\vector(0,-1){35}}
\put (20,63){\makebox(0,0){Chained record}}
\put (80,63){\makebox(0,0){Unchained record}}
\put (42,47){(4 bytes)}
\put (102,47){(237 bytes)}
\put (44,22){FRAME containing domain D}
\put (44,17){(3519 bytes)}
\put (44,12){(7 PDBs)}
\end {picture}
\caption {An Example of a Chained and an Unchained Record}
\label {an_example_of_a_chained_and_an_unchained_record}
\end {center}
\end {figure}

The unchained record's domain D contains 237 bytes of data which can be stored
directly within the confines of the LRB. The chained record's domain D length
of 3519 bytes results in the actual data being stored in a frame of seven
contiguous PDBs; the D domain in the LRB becomes a 4-byte field containing a
pointer to the first block in the frame.

\subsubsection {Container and Component Records}

These records are similar in function. A Container Record can be regarded as a
top level Component Record. It is the first record in every container file,
ie.\ it starts in the first chip of the second block. Their structures are
illustrated in Figure~\ref{structure_of_container_and_component_records}.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(100,40)
\thicklines
\put (0,30){\framebox(40,5){RCL}}
\put (0,25){\framebox(40,5){null}}
\put (0,20){\framebox(40,5){top-level name}}
\put (0,15){\framebox(40,5){RID}}
\put (60,30){\framebox(40,5){RCL}}
\put (60,25){\framebox(40,5){no.\ of components}}
\put (60,20){\framebox(40,5){component name 1}}
\put (60,15){\framebox(40,5){RID}}
\put (60,5){\framebox(40,5){component name n}}
\put (60,0){\framebox(40,5){RID}}
\put (112,32){Domain C}
\put (112,27){Domain S}
\put (112,22){Domain D}
\put (20,38){\makebox(0,0){{\bf Container}}}
\put (80,38){\makebox(0,0){{\bf Component}}}
\put (42,32){(10)}
\put (42,27){(0)}
\put (42,22){(16)}
\put (42,17){(4)}
\put (102,32){(10)}
\put (102,27){(4)}
\put (102,22){(16)}
\put (102,17){(4)}
\put (102,7){(16)}
\put (102,2){(4)}
\put (60,15){\line(0,-1){1}}
\put (60,13){\line(0,-1){1}}
\put (60,11){\line(0,-1){1}}
\put (100,15){\line(0,-1){1}}
\put (100,13){\line(0,-1){1}}
\put (100,11){\line(0,-1){1}}
\put (0,30.5){\line(1,0){40}}
\put (0,25.5){\line(1,0){40}}
\put (0,15.5){\line(1,0){40}}
\put (60,30.5){\line(1,0){40}}
\put (60,25.5){\line(1,0){40}}
\put (60,0.5){\line(1,0){40}}
\end {picture}
\caption {Structure of Container and Component Records}
\label {structure_of_container_and_component_records}
\end {center}
\end {figure}

A Component Record is basically a list of object names and associated RIDs. If a
component is a structure, its RID will point to a Structure Record; if it is a
primitive, the RID will point to a Primitive Record. Consider a case where a
structure has three components, RA, DEC and EQUINOX. Its component record's
static domain contains the value 3 and the dynamic domain is illustrated in
Figure~\ref{an_example_of_the_dynamic_domain_in_a_component_record}.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(120,10)
\thicklines
\put (0,5){\framebox(35,5){RA}}
\put (35,5){\framebox(5,5){*}}
\put (40,5){\framebox(35,5){DEC}}
\put (75,5){\framebox(5,5){*}}
\put (80,5){\framebox(35,5){EQUINOX}}
\put (115,5){\framebox(5,5){*}}
\put (37.5,7.5){\vector(0,-1){7.5}}
\put (77.5,7.5){\vector(0,-1){7.5}}
\put (117.5,7.5){\vector(0,-1){7.5}}
\end {picture}
\caption {An Example of the Dynamic Domain in a Component Record}
\label {an_example_of_the_dynamic_domain_in_a_component_record}
\end {center}
\end {figure}

\subsubsection {Structure and Primitive Records}

These records are collectively known as {\em Object Records}. They are similar
in function and their structures are illustrated in
Figure~\ref{structure_of_structure_and_primitive_records}.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(100,25)
\thicklines
\put (0,20){\framebox(40,5){RCL}}
\put (0,15){\framebox(40,5){ODL}}
\put (0,10){\framebox(40,5){RID}}
\put (0,0){\framebox(40,5){RID}}
\put (60,20){\framebox(40,5){RCL}}
\put (60,15){\framebox(40,5){ODL}}
\put (60,10){\framebox(40,5){value}}
\put (60,0){\framebox(40,5){value}}
\put (112,22){Domain C}
\put (112,17){Domain S}
\put (112,12){Domain D}
\put (20,28){\makebox(0,0){{\bf Structure}}}
\put (80,28){\makebox(0,0){{\bf Primitive}}}
\put (42,22){(10)}
\put (42,12){(4)}
\put (42,2){(4)}
\put (102,22){(10)}
\put (0,10){\line(0,-1){1}}
\put (0,8){\line(0,-1){1}}
\put (0,6){\line(0,-1){1}}
\put (40,10){\line(0,-1){1}}
\put (40,8){\line(0,-1){1}}
\put (40,6){\line(0,-1){1}}
\put (60,10){\line(0,-1){1}}
\put (60,8){\line(0,-1){1}}
\put (60,6){\line(0,-1){1}}
\put (100,10){\line(0,-1){1}}
\put (100,8){\line(0,-1){1}}
\put (100,6){\line(0,-1){1}}
\put (0,20.5){\line(1,0){40}}
\put (0,15.5){\line(1,0){40}}
\put (0,0.5){\line(1,0){40}}
\put (60,20.5){\line(1,0){40}}
\put (60,15.5){\line(1,0){40}}
\put (60,0.5){\line(1,0){40}}
\end {picture}
\caption {Structure of Structure and Primitive Records}
\label {structure_of_structure_and_primitive_records}
\end {center}
\end {figure}

The static domain contains an {\em Object Descriptor Label} (ODL), which
is illustrated in Figure~\ref{structure_of_an_object_descriptor_label}.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,25)
\thicklines
\put (0,0){\framebox(40,5){axis [naxes]}}
\put (0,10){\framebox(40,5){axis [2]}}
\put (0,15){\framebox(40,5){axis [1]}}
\put (0,20){\framebox(10,5){naxes}}
\put (10,20){\framebox(30,5){type}}
\put (0,10){\line(0,-1){1}}
\put (0,8){\line(0,-1){1}}
\put (0,6){\line(0,-1){1}}
\put (40,10){\line(0,-1){1}}
\put (40,8){\line(0,-1){1}}
\put (40,6){\line(0,-1){1}}
\put (42,22){($1+15$ bytes)}
\put (42,17){(4 bytes)}
\put (42,12){(4 bytes)}
\put (42,2){(4 bytes)}
\end {picture}
\caption {Structure of an Object Descriptor Label}
\label {structure_of_an_object_descriptor_label}
\end {center}
\end {figure}

Primitive Records can be distinguished from Structure Records by the fact that
the TYPE specification in the ODL corresponds to one of the primitive data
types. Figure~\ref{examples_of_object_descriptor_labels} illustrates some
examples of ODLs.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(50,55)
\thicklines
\put (0,45){\framebox(40,5){equatorial}}
\put (40,45){\framebox(5,5){0}}
\put (0,30){\framebox(40,5){spectrum}}
\put (40,30){\framebox(5,5){2}}
\put (45,30){\framebox(10,5){8}}
\put (55,30){\framebox(10,5){9}}
\put (0,15){\framebox(40,5){\_real}}
\put (40,15){\framebox(5,5){2}}
\put (45,15){\framebox(10,5){512}}
\put (55,15){\framebox(10,5){256}}
\put (0,0){\framebox(40,5){\_char*80}}
\put (40,0){\framebox(5,5){1}}
\put (45,0){\framebox(10,5){36}}
\put (0,53){A scalar structure of type `equatorial'}
\put (0,38){An $8 \times 9$ structure array of type `spectrum'}
\put (0,23){A $512 \times 256$ single precision floating point array}
\put (0,8){A 36-element character vector}
\end {picture}
\caption {Examples of Object Descriptor Labels}
\label {examples_of_object_descriptor_labels}
\end {center}
\end {figure}

\subsection {Manipulation of Records}

The following sections describe the processes by which various types of record
manipulation are implemented.

\subsubsection {Record access}

Records are accessed by the DAT module through their RIDs. However, since a
program can have more than one file open, further qualification is required and
the RID and file ({\em slot} --- see Section~\ref{access_to_files}) number
together are used uniquely to identify a record. This entity is termed a {\em
handle} and its structure is illustrated in Figure~\ref{structure_of_a_handle}.
It is analogous to the {\em locator} used in application programs to identify
an object. The relationship is illustrated in
Figure~\ref{entities_handled_by_the_dat_and_rec_modules}.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,10)
\thicklines
\put (0,5){\framebox(40,5){RID}}
\put (0,0){\framebox(10,5){}}
\put (10,0){\framebox(10,5){read}}
\put (20,0){\framebox(20,5){slot}}
\put (42,7){(4 bytes)}
\put (42,2){($1+1+6$ bits)}
\end {picture}
\caption {Structure of a Handle}
\label {structure_of_a_handle}
\end {center}
\end {figure}

\begin {figure}[htbp]
\begin {center}
\begin {picture}(85,23)
\thicklines
\put (24,20){{\bf DAT module}}
\put (64,20){{\bf REC module}}
\put (0,10){{\bf entity:}}
\put (0,5){{\bf address:}}
\put (30,10){object}
\put (30,5){locator}
\put (70,10){record}
\put (70,5){handle}
\put (45,11){\vector(1,0){20}}
\put (65,11){\vector(-1,0){20}}
\put (45,6){\vector(1,0){20}}
\put (65,6){\vector(-1,0){20}}
\end {picture}
\caption {Entities Handled by the DAT and REC Modules}
\label {entities_handled_by_the_dat_and_rec_modules}
\end {center}
\end {figure}

\subsubsection {Record creation}

Because of the hierarchical nature of the system, the DAT module must supply a
handle to the parent record when invoking the record creation routine. To
minimise the processing overheads, the strategy is to try first to create the
new record in the same LRB as its parent. If this fails the free space stack is
searched for an LRB with enough contiguous free chips to satisfy the request.
If this fails a virgin LRB must be found, either by grabbing a free block or by
extending the file if necessary. A frame of PDBs must also be allocated if the
record's dynamic domain has to be chained due to its size.

\subsubsection {Record deletion}

If the dynamic domain is chained, the associated frame of PDBs is first
returned to the free space stack. The appropriate bits in the LRB's chip bitmap
are then cleared (indicating that the chips are free) and finally the free
space stack is updated to reflect the deallocation of the record's chips.

\subsubsection {Record expansion}

The expansion of an existing record's dynamic domain is probably the most
complex operation performed by the record management facility. In the simplest
cases the required extra space can be found at the current end of the domain
--- if the record is chained the last block in the PDB frame may have
sufficient space to accommodate the expansion; if unchained there may be room
in the last chip, or there may be enough free chips adjacent to the current end
of record to satisfy the request. If there is insufficient space to directly
expand an unchained record the dynamic domain must be chained --- a PDB frame
of appropriate size is first allocated from the free space stack, the current
contents of the domain moved and then overwritten with the block number of the
new frame. Because this only requires 4 bytes the record's dynamic domain
contracts which may result in the deallocation of chips. If the record is
already chained the frame of PDBs must be expanded. Depending on the current
state of the file, one of three actions will be performed:

\begin {itemize}
\item If there are sufficient contiguous free blocks after the end of the frame,
these are allocated.
\item If all the blocks between the end of the frame and the end of the file are
free, but there aren't enough to satisfy the request, the file is extended.
\item In the worst case where the frame cannot be directly expanded, a new frame
of the required size must be allocated and the existing data moved.
The old frame is returned to the free space stack.
\end {itemize}

\subsubsection {Record contraction}

If the domain is chained, contraction can result in PDBs being returned to the
free space stack. If chips are released from an unchained record the chip
bitmap is updated.

\subsection {Access to Files}
\label {access_to_files}

Container files are opened by the RMS service {\tt \$OPEN} with the {\em User
File Open} (UFO) option which allows a process to take advantage of the
file-mapping facilities provided by VMS. Furthermore, as this service returns an
I/O channel to the file the {\tt \$QIO} System Service can also be utilised in
those situations where mapping is expensive. VMS restricts the number of mapped
sections a process can have to a typical limit of 32. The {\em Create and Map
Section} system service {\tt \$CRMPSC} is only used to access PDBs; LRBs and
HCBs are read and written using {\tt \$QIO}s. For details of VAX/VMS System
Services see \cite{system_services_reference}.

To reduce I/O activity, LRBs and HCBs are cached on a {\em Working Page List}
--- 16 Kbytes of cache space is provided which allows up to 32 blocks to reside
in physical memory at any instant. When the list becomes full and a new block
is required, an existing block must be released. If the block has been modified
it is written back to the file before the physical memory space is re-used.
Each block is controlled by an associated {\em Block Control Packet} (BCP).
The structure of a BCP is illustrated by
Figure~\ref{structure_of_a_block_control_packet}.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,5)
\thicklines
\put (0,0){\framebox(10,5){}}
\put (10,0){\framebox(10,5){slot}}
\put (20,0){\framebox(20,5){block}}
\put (42,2){($6+6+20$ bits)}
\end {picture}
\caption {Structure of a Block-ID}
\label {structure_of_a_block_id}
\end {center}
\end {figure}

\begin {figure}[htbp]
\begin {center}
\begin {picture}(50,25)
\thicklines
\put (0,20){\framebox(50,5){pointer to next BCP}}
\put (0,15){\framebox(50,5){pointer to last BCP}}
\put (0,10){\framebox(50,5){pointer to block in memory}}
\put (0,5){\framebox(50,5){block-id}}
\put (0,0){\framebox(5,5){}}
\put (5,0){\framebox(15,5){modified}}
\put (20,0){\framebox(30,5){reference count}}
\put (52,22){(4 bytes)}
\put (52,17){(4 bytes)}
\put (52,12){(4 bytes)}
\put (52,7){(4 bytes)}
\put (52,2){($2+1+5$ bits)}
\end {picture}
\caption {Structure of a Block Control Packet}
\label {structure_of_a_block_control_packet}
\end {center}
\end {figure}

When a file is first opened (or created) it is assigned a {\em slot} in the
{\em File Control Vector} (FCV). An FCV slot is illustrated in
Figure~\ref{structure_of_a_file_control_vector_slot}. Each active slot contains
such information as the expanded file-name string, device, directory and
file-IDs, read-only and read/write I/O channels, and various bit flags which
describe the status and disposition of the file. A reference count is kept of
the number of explicit open operations performed by calls to {\tt HDS\_OPEN}; a
file is not physically closed until this count drops to zero. Slots in the FCV
are also allocated to the {\em temporary} container files which are created in
the user's {\tt SYS\$SCRATCH} directory and automatically deleted when closed.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,15)
\thicklines
\put (0,10){\framebox(40,5){device id}}
\put (0,5){\framebox(40,5){file id}}
\put (0,0){\framebox(40,5){directory id}}
\put (42,12){($1+15$ bytes)}
\put (42,7){($3 \times 2$ bytes)}
\put (42,2){($3 \times 2$ bytes)}
\end {picture}
\caption {Structure of a File-ID}
\label {structure_of_a_file_id}
\end {center}
\end {figure}

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,35)
\thicklines
\put (0,30){\framebox(40,5){pointer to filename}}
\put (0,25){\framebox(40,5){pointer to file-id}}
\put (0,20){\framebox(40,5){lock id}}
\put (0,15){\framebox(40,5){read channel}}
\put (0,10){\framebox(40,5){write channel}}
\put (0,5){\framebox(40,5){reference count}}
\put (0,0){\framebox(16,5){reserve}}
\put (16,0){\framebox(8,5){lock}}
\put (24,0){\framebox(8,5){open}}
\put (32,0){\framebox(8,5){dele}}
\put (42,32){(4 bytes)}
\put (42,27){(4 bytes)}
\put (42,22){(4 bytes)}
\put (42,17){(2 bytes)}
\put (42,12){(2 bytes)}
\put (42,7){(2 bytes)}
\put (42,2){($13+1+1+1$ bits)}
\end {picture}
\caption {Structure of a File Control Vector Slot}
\label {structure_of_a_file_control_vector_slot}
\end {center}
\end {figure}

\subsection {Sharing of Files}

Container files can be accessed concurrently by multiple processes and
subprocesses. Read-only sharing in a multi-user environment generally presents
few, if any, real problems and this facility is provided free by HDS. However,
processing costs can rise sharply if multiple concurrent writers are allowed,
particularly under a local buffer caching scheme such as that employed by HDS.
The system does, in fact, support concurrent writers, but on a synchronous
rather than asynchronous basis, ie.\ although a file may be open for
write-access by multiple processes, only one will be able to actually perform
write operations at any given time. This is achieved through the use of the
{\tt \$ENQ} System Service which is used to grant a process exclusive
write-access to a specific container file, either explicitly by a call to {\tt
HDS\_LOCK} or automatically when the process performs its first write
operation. If the file is currently locked by another process, the requesting
process has the option of waiting or receiving control back immediately; refer
to the section on tuning for more information. A process can relinquish a lock
on a container file by calling {\tt HDS\_FREE}. This writes back all modified
cached blocks to the file and then releases the {\tt \$ENQ} lock thereby
granting write-access to other processes. HDS rundown automatically releases
all container file locks\footnote{A call to {\tt HDS\_FREE} does {\em not}
unmap data. This is potentially a serious problem.}.

\subsection {Management of Disc Space}
\label {disc_space_management}

The {\em Free Space Stack} is a part of the HCB. It grows from both ends like
Process Space on the VAX. Entries at the bottom describe which  PDB frames are
free; entries at the top describe the largest free contiguous space in each LRB.

Consider the configuration illustrated in
Figure~\ref{an_example_of_a_free_space_stack}. Examining the bottom of the
stack, there are 17 free blocks starting at block 101 and 3 spare at 8; these
blocks are available as either new PDB frames or individual LRBs. The top of the
stack contains three entries: LRB 2 has at most 5 contiguous free chips (see
below), LRB 7 has none, and 19 has 10 contiguous chips as its largest free
space. The stack is initially filled with {\em null} entries (-1) which
gradually disappear as the opposing ends of the stack grow towards each other.
When only one null entry is left the stack is regarded as full and no new
entries can be added (although {\em void} entries, ie.\ those with zero free
space, can be re-used). This gradual degradation of a file allows tidying-up to
be deferred until the free space stack is full rather than incorporating this
function into normal application processing.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,40)
\thicklines
\put (0,0){\framebox(20,5){17}}
\put (0,5){\framebox(20,5){3}}
\put (0,10){\framebox(20,5){-1}}
\put (0,20){\framebox(20,5){-1}}
\put (0,25){\framebox(20,5){10}}
\put (0,30){\framebox(20,5){0}}
\put (0,35){\framebox(20,5){5}}
\put (20,0){\framebox(20,5){101}}
\put (20,5){\framebox(20,5){8}}
\put (20,10){\framebox(20,5){-1}}
\put (20,20){\framebox(20,5){-1}}
\put (20,25){\framebox(20,5){19}}
\put (20,30){\framebox(20,5){7}}
\put (20,35){\framebox(20,5){2}}
\put (0,20){\line(0,-1){1}}
\put (0,18){\line(0,-1){1}}
\put (0,16){\line(0,-1){1}}
\put (40,20){\line(0,-1){1}}
\put (40,18){\line(0,-1){1}}
\put (40,16){\line(0,-1){1}}
\put (42,40){\vector(0,-1){15}}
\put (42,25){\vector(0,1){15}}
\put (42,10){\vector(0,-1){10}}
\put (42,0){\vector(0,1){10}}
\put (44,32){LRB space}
\put (44,4){PDB space}
\end {picture}
\caption {An Example of a Free Space Stack}
\label {an_example_of_a_free_space_stack}
\end {center}
\end {figure}

When the free space stack becomes full any space subsequently released will
effectively be lost. When a file has degraded this far it may be advisable to
perform a tidying-up operation such as creating a new container file and
copying all the objects from the existing file, preserving the logical
structure.

\subsection {Management of Dynamic Memory}

The REC module supervises the dynamic allocation of virtual memory within
process address space. When REC is initialised the program region is expanded
to accommodate a bitmap which is used subsequently to reflect the state of
virtual memory pages (ie.\ whether they are allocated or free). The size of
this bitmap is calculated from the VMS system parameter VIRTUALPAGECNT which
specifies the maximum number of virtual pages that can be mapped for any
process. The VMS System Symbol Table {\tt SYS\$SYSTEM:SYS.STB} must be scanned
by the VAX Linker to resolve the address of this parameter.

Access to the bitmap and the virtual memory controlled by it is via a {\em
Virtual Memory Pool} (VMP) structure which is illustrated in
Figure~\ref{structure_of_the_virtual_memory_pool}. Initially the bitmap
describes only itself.

When large amounts of memory are required or a PDB frame is to be mapped from a
container file, the bitmap is searched on a first-fit basis for a contiguous
sequence of free virtual memory pages. If found, they are allocated and the
appropriate bits in the bitmap set. If there is insufficient space in the
current program region it must be expanded.

When small amounts of memory are required the VMS RTL routine {\tt LIB\$GET\_VM}
is invoked. This may itself result in an expansion of the program region. If
this happens then the next time that the HDS memory management routines expand
the program region the bits will be set in the bitmap corresponding both to the
pages got earlier using {\tt LIB\$GET\_VM} and to the pages just got. The
bits corresponding to the memory allocated using {\tt LIB\$GET\_VM} will
never get reset, so HDS will never attempt to use that memory.

\begin {figure}[htbp]
\begin {center}
\begin {picture}(40,15)
\thicklines
\put (0,10){\framebox(40,5){pointer to bitmap}}
\put (0,5){\framebox(40,5){pool start address}}
\put (0,0){\framebox(40,5){pool end address}}
\put (42,12){(4 bytes)}
\put (42,7){(4 bytes)}
\put (42,2){(4 bytes)}
\end {picture}
\caption {Structure of the Virtual Memory Pool}
\label {structure_of_the_virtual_memory_pool}
\end {center}
\end {figure}

All data in LRBs is accessed via direct I/O. All data in PDBs is mapped. This
means that it is only an object's {\em size} that governs the method of access
--- the largest object that can possibly fit into an LRB has a size of 480
bytes (an RCL plus the ODL of a 1-D object occupy $10+16+4$ bytes which leaves
$510-30$ or 480 bytes for the data) so all larger objects will be mapped
regardless of whether they are being accessed using {\tt DAT\_GET}, {\tt
DAT\_MAP} or {\tt DAT\_BASIC}.

A separate question is whether the application receives a direct pointer
to or copy of the object data in the container file. This can only happen
if the data is contiguous and if no type conversion is necessary. The most
efficient way to access data is thus to use {\tt DAT\_MAP} or {\tt DAT\_BASIC},
to use contiguous slices and to access the data in the same format as it
is stored on disc. If you only intend accessing a part of a large data object
then it is well worthwhile using {\tt DAT\_SLICE} to obtain a locator to
the bit in which you are interested. Not only does this save virtual memory,
but it saves the system from having to dispose of that virtual memory.

\section {HDS IMPLEMENTATION}

Both the DAT and REC modules have a set of externally-visible routines and a set
of internal routines. DAT's external routines are those which are called by
applications. REC's external routines are those which are called by DAT. (The
EXC routines may be called by the application, by DAT and by REC).
Appendix~\ref{routine_lists} contains reference lists of all the routines
in all the modules.

This section explains source conventions and rules with the specific aim of
allowing easy understanding of high-level DAT routines and safe implementation
of new such routines.

\subsection {Use of Global Variables}
\label {use_of_global_variables}

HDS makes very limited use of global variables. Those that are used fall
into three categories:

\begin {itemize}
\item A status variable that is shared by all modules.
\item Current values for the quantities that can be tuned --- see
Section~\ref{system_tuning}.
\item Miscellaneous quantities that is is convenient to have accessible
in other modules. There are not many of these.
\end {itemize}

The global status variable is called {\tt HDS\_GL\_STATUS}. It is used in
the following manner:

\begin {itemize}
\item On entry to all user-callable DAT routines it is set to the HDS success
status {\tt DAT\_\_OKAY}.
\item Any routine that fails simply sets {\tt HDS\_GL\_STATUS} and this
is frequently checked using a set of macros --- see
Section~\ref{use_of_macros}.
\end {itemize}

\subsection {Use of Macros}
\label {use_of_macros}

Several utility macros are used throughout HDS. Nearly all of them have
names that begin with underscore characters to allow easy identification.
They split up into the following categories:

\begin {description}
\item [Simple utility functions]: calculate the {\tt min} and {\tt max} of two
numbers --- these are the only macros whose names do not begin with underscores.
\item [Moving data around]: these all begin with the characters {\tt \_ch}
and are all wraps around calls to VMS or C RTL routines.
\item [Accessing system calls]: {\tt \_getjpiw}, {\tt \_dclexh}, {\tt \_canexh}
and {\tt \_valrms} all access system services or RMS routines.
\item [Common expression evaluation]: {\tt \_nchips} and {\_nblocs} perform
common calculations.
\item [Avoiding call overheads]: {\tt \_seek\_bloc} and {\tt \_drop\_bloc}
avoid what would otherwise be subroutine calls.
\item [Checking status]: {\tt \_ok} checks whether a status value is an
OK HDS status.
\item [Entering and exiting routines]: these all begin {\tt \_enter} or
{\tt \_exit} and are described below.
\item [Invoking other routines]: {\tt \_invoke} and {\tt \_call} are described
below.
\end {description}

\subsubsection {Entering, exiting and calling routines}

Every HDS routine begins with a line of the form:

\begin {quote}
\begin {verbatim}
_enter(ROUTINE_NAME)
\end{verbatim}
\end {quote}

This is to permit keeping track of how many times HDS routines have been
called. Normally the {\tt \_enter} macro has an empty definition.

When an HDS routine invokes another routine it has three choices:

\begin {itemize}
\item It can call the routine directly,
\item It can use the {\tt \_invoke} macro, or
\item It can use the {\tt \_call} macro.
\end {itemize}

If the routine is called directly then any error checking must be explicit.

The {\tt \_invoke (proc)} macro looks like:

\begin {quote}
\begin {verbatim}
{
proc;
if (!_ok(hds_gl_status)) _exit0;
}
\end{verbatim}
\end {quote}

In other words, it invokes the routine and then, if the global status is
not still OK, it exits using the {\tt \_exit0} macro, which simply does:

\begin {quote}
\begin {verbatim}
return hds_gl_status;
\end{verbatim}
\end {quote}

The {\tt \_call (event)} macro is similar but slightly more complicated. It
looks like:

\begin {quote}
\begin {verbatim}
{
*status = (event);
if (!_ok(*status))
    _exit1(exc_$hds(status));
}
\end{verbatim}
\end {quote}

It can be seen that it is designed for use in routines that have a {\tt STATUS}
variable and for calling routines that return a status as their function value
as well as or instead of setting {\tt HDS\_GL\_STATUS}. If the called routine
returns a bad status, the HDS exception handler {\tt EXC\_\$HDS} is called to
handle and / or report that error. The default behaviour of {\tt EXC\_\$HDS} is
to return the input status as its function result, so this bad status will now
be returned to the calling routine via the {\tt \_exit1 (status)} macro, which
simply does:

\begin {quote}
\begin {verbatim}
{
hds_gl_status = status;
return hds_gl_status;
}
\end{verbatim}
\end {quote}

{\tt \_call} is only used in the top-level DAT routines, all of which have {\tt
STATUS} parameters. {\tt \_invoke} is used in low-level DAT routines and all
REC routines, all of which return status as their function values.

The {\tt \_exit0} and {\tt \_exit1} macros have already been introduced.
{\tt \_exit0} is used when {\tt HDS\_GL\_STATUS} has already got the value
which is to be returned as the function value. {\tt \_exit1} is the same
except that it takes a {\tt STATUS} and copies it to {\tt HDS\_GL\_STATUS}
before returning it. There are also {\tt \_exit2} and {\tt \_exit3} macros
which are less relevant to this discussion.

Examination of the macro definitions shows that the following forms are
both valid:

\begin {minipage}[t]{80mm}
\begin {quote}
\begin {verbatim}
_call (DAT__xxxxx)
\end{verbatim}
\end {quote}
\end {minipage}
\begin {minipage}[t]{80mm}
\begin {quote}
\begin {verbatim}
_exit1 (DAT_xxxxx)
\end{verbatim}
\end {quote}
\end {minipage}

The former is very commonly used in the high-level DAT routines as a means of
returning a failure status to an application routine and the latter is very
commonly used in the low-level DAT routines and the REC routines for returning
a failure status to a high-level DAT routine.

Finally, {\em don't} put semi-colons after any of these macro calls! They
are not required since all of them are blocks rather than single statements.
Often, such a semi-colon will do no harm, but it can do in a case such
as this:

\begin {minipage}[t]{80mm}
\begin {quote}
\begin {verbatim}
if (a == b)
    _call (DAT__OBJIN);





else
    printf ("not equal");
\end{verbatim}
\end {quote}
\end {minipage}
\begin {minipage}[t]{80mm}
\begin {quote}
\begin {verbatim}
if (a == b)
    {
    *status = DAT__OBJIN;
    if (!_ok(*status))
        _exit1(exc_$hds(status));
    }
    ;
else
    printf ("not equal");
\end{verbatim}
\end {quote}
\end {minipage}

\subsection {Use of Exception Handlers}

When an error occurs, HDS always calls {\tt EXC\_\$VMS} or {\tt EXC\_\$HDS}
just before returning to the calling application. The idea is that {\tt
EXC\_\$HDS} is called for HDS errors and {\tt EXC\_\$VMS} is called for VMS
errors.

In the VAX/VMS implementation these two routines are identical --- {\tt
EXC\_\$HDS} simply calls {\tt EXC\_\$VMS} --- because there is no overlap
between HDS and VMS error codes. If HDS were ever implemented on a system where
this was not possible then the two routines would have to be different.
If this happened it would also be necessary to make {\tt EXC\_\$VMS} return
an HDS error code as its function value since otherwise high-level DAT routines
would be able to return ambiguous status values (this is what used to happen
but it had the unfortunate side-effect that if a VMS error occurred there
was no indication of what exactly had gone wrong).

All {\tt EXC\_\$HDS} calls are made by the {\tt \_call} macro from one of
the high-level DAT routines. All {\tt EXC\_\$VMS} calls are made directly
from the REC routines. When a {\tt EXC\_\$VMS} call is made from REC there
will also be a later {\tt EXC\_\$HDS} call from the DAT routines, which
will result in such errors being reported twice. This is a consequence of
the fact that {\tt EXC\_\$VMS} doesn't return an HDS error code --- see
the discussion of the previous paragraph.

\subsection {Inter-module Interfaces}

The use of global variables is described in
Section~\ref{use_of_global_variables}. Not surprisingly, apart from the use of
global variables, all the communication between the DAT and REC modules is via
argument lists. Nearly all such arguments are simple scalar variables, although
there are a few structures.

Table~\ref{argument_names_used_in_the_high_level_rec_routines} is a list
ordered by type of argument names used in the high-level REC routines.
Apparently HAN, RID and RCL are the only structures whose definitions are
known by both DAT and REC. In fact this is not the case because DAT regards
both HAN and RID as being integers. Thus RCL is the only structure whose
definition is known to both modules. Typically DAT sets up some of its fields
and REC fills in the rest of them.

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|l|l||}
\hline
Data type	& Passing Mechanism	& Argument names \\
\hline
{\tt INTEGER}	& By value		& {\tt EXTENT,LENGTH,OFFSET,SIZE} \\
{\tt CHARACTER}	& Address of descriptor	& {\tt FILE} \\
{\tt CHAR}	& By value		& {\tt STATE,MODE,DISP} \\
{\tt CHAR}	& By reference		& {\tt PNTR,BUF} \\
{\tt CHAR}	& Pointer by reference	& {\tt FNS} \\
\hline
{\tt HAN}	& By reference		& {\tt HAN,HAN1,HAN2,PAR,KIN,SRC,DES} \\
{\tt RID}	& By reference		& {\tt RID} \\
{\tt RCL}	& By reference		& {\tt RCL} \\
\hline
\end {tabular}
\caption {Argument Names Used in the High-level REC Routines}
\label {argument_names_used_in_the_high_level_rec_routines}
\end {center}
\end {table}

\subsection {Use of Locators}
\label {use_of_locators}

Section~\ref{locators} introduced the LCP and indicated that it is used
for caching object information that would otherwise be held only on disc.
In this section, the structure and usage of the LCP are described in detail.

\subsubsection {Structure of the LCP}

The LCP structure is illustrated in
Figure~\ref{structure_of_a_locator_control_packet}. The three flags {\bf valid},
{\bf struc} and {\bf read} signify the following when set:

\begin {description}
\item [valid]: the fields of the LCP are valid (when the LCP is first allocated
all its fields are zero and it is therefore not valid)
\item [struc]: the LCP describes a structure object (ie.\ the first character
of the object type is not an underscore character)
\item [read]: only read access is permitted to the object described by the
LCP (this is always inherited from the object's parent and is therefore
the same for all objects in a given container file)
\end {description}

\begin {figure}[htbp]
\begin {center}
\begin {picture}(64,75)
\thicklines
\put (0,0){\framebox(64,5){object data descriptor}}\put (66,2){(8 bytes)}
\put (0,5){\framebox(64,5){application data descriptor}}\put (66,7){(8 bytes)}
\put (0,10){\framebox(64,5){dimension bounds $(l_{1},u_{1},l_{2},u_{2},l_{3},%
u_{3})$}}\put (66,12){($6 \times 4$ bytes)}
\put (0,15){\framebox(16,5){naxes}}
\put (16,15){\framebox(16,5){mode}}
\put (32,15){\framebox(32,5){level}}\put (66,17){($1+1+2$ bytes)}
\put (0,20){\framebox(64,5){offset}}\put (66,22){(4 bytes)}
\put (0,25){\framebox(64,5){RID of parent component record}}\put (66,27){(4 %
bytes)}
\put (0,30){\framebox(64,5){handle to object record}}\put (66,32){(4 bytes)}
\put (0,35){\framebox(64,5){product of object dimensions}}\put (66,37){(4 %
bytes)}
\put (0,40){\framebox(4,5){}}
\put (4,40){\framebox(60,5){object name}}\put (66,42){($1+15$ bytes)}
\put (0,45){\framebox(4,5){}}
\put (4,45){\framebox(60,5){object type}}\put (66,47){($1+15$ bytes)}
\put (0,50){\framebox(4,5){}}
\put (4,50){\framebox(60,5){group name}}\put (66,52){($1+15$ bytes)}
\put (0,55){\framebox(32,5){reserve}}
\put (32,55){\framebox(32,5){object state}}\put (66,57){($8+8 \times 1$ bits)}
\put (0,60){\framebox(16,5){reserve}}
\put (16,60){\framebox(16,5){read}}
\put (32,60){\framebox(16,5){struc}}
\put (48,60){\framebox(16,5){valid}}\put (66,62){($13+3 \times 1$ bits)}
\put (0,65){\framebox(64,5){link to previous LCP}}\put (66,67){(4 bytes)}
\put (0,70){\framebox(64,5){link to next LCP}}\put (66,72){(4 bytes)}
\end {picture}
\caption {Structure of a Locator Control Packet}
\label {structure_of_a_locator_control_packet}
\end {center}
\end {figure}

There are seven flags describing the object state. Actually it is more correct
to say that they describe the view presented by the LCP of the object because
the same object can simultaneously be viewed in several different ways using
different LCPs. The flags are:

\begin {description}
\item [mapped]: object data (ie.\ the object record's dynamic domain) is
currently mapped into virtual memory by the application
\item [vmcopy]: the application is mapped to a virtual memory copy of the
object data. This is only used when the {\bf mapped} flag is set. It can
be set either because type conversion was necessary or because the object's
dynamic domain is not chained (so it is in an LRB rather than a PDB)
\item [unlike]: the object data type is different from that requested by
the application and so type conversion was necessary. This is only used when
the {\bf mapped} flag is set
\item [slice]: the LCP describes a slice of the object. A slice is defined by
sub-ranges in up to the first three dimensions
\item [cell]: the LCP describes a single cell (or element) of the object
\item [vector]: the LCP describes a vectorised object. Regardless of the
actual dimensionality of the object, it is considered to be a 1-D array when
accessed using this LCP
\item [broken]: the LCP describes a non-contiguous part of the object. This
can only be the case when the {\bf slice} flag is set as well
\end {description}

A few of the other LCP fields may require a little explanation:

\begin {description}
\item [group]: is a character string that can be used to associate related
LCPs. It can be set using {\tt HDS\_LINK} and then is inherited by all
lower-level LCPs. {\tt HDS\_FLUSH} can be used to unmap all data associated
with all LCPs with a given group specification and then to mark those LCPs
as being invalid
\item [offset and dimension bounds]: are used for describing slices, cells
and vectorised objects
\item [mode]: is only used for LCPs associated with mapped data - it is
the access mode (Read / Write / Update) with which the data is mapped
\item [level]: is the level of this object within the container file. A
top-level object is at level zero
\item [data descriptors]: VMS descriptors are used to describe both the
application's view of the data and the object's view of the data. The class,
type and length fields are filled in in the object data descriptor when
the LCP is created. The pointer fields and all the fields in the application
data descriptor are filled in when the data is read, written or mapped.
Many of the internal routines, eg.\ the type conversion routines, use these
descriptors
\end {description}

\subsubsection {Administration of the LCP}

When a new locator is to be created, an {\em export locator} routine is called.
This allocates and zeroes a new LCP, then arranges for the locator to point
to this LCP. The result of this is to create a new and invalid locator.
The high-level DAT routine will then fill in fields --- normally the last
field to be filled in is the {\bf valid} flag.

When an existing locator is to be used, an {\em import locator} routine
is called. This routine performs several checks in an attempt to ensure
that attempts to use invalid locators will be detected:

\begin {itemize}
\item the locator variable passed from the application is indeed {\tt
CHARACTER*15}
\item the validity check field is correct
\item the LCP {\bf valid} flag is set
\item the parent of the record whose handle is in the LCP is the same as
the parent field in the LCP
\end {itemize}

When an existing locator is to be de-activated, the LCP valid flag is reset
and the LCP is put on a free list. A future attempt to use this LCP will
fail unless the same LCP has by chance subsequently been associated with
exactly the same object.

\subsection {Structure of a High-level DAT Routine}

Most of the high-level DAT routines have a fairly standard structure:

\begin {itemize}
\item exit if status is bad and set global status to {\tt DAT\_\_OKAY}
\item import any input locators
\item check LCP fields to see if the proposed operation is valid
\item validate other arguments
\item export any output locators
\item convert fields from input locators to output locators
\item perform operation-specific processing
\item set output locator state flags
\end {itemize}

Consider a genuine but not too long example: {\tt DAT\_GET}. Perhaps
surprisingly, {\tt DAT\_GET}, {\tt DAT\_PUT} and {\tt DAT\_MAP} are amongst
the most straightforward of the routines. The code in what follows is lifted
directly from the current version of the C code. The only changes that have
been made is that some declarations have been removed and the comments
have been enhanced and typeset.

\begin {quote}
\begin {verbatim}
dat_get(locator,type,ndim,dims,values,status) {
\end{verbatim}
\end {quote}

Enter routine and import the locator.

\begin {quote}
\begin {verbatim}
if (!_ok(*status))
        return;
hds_gl_status      = DAT__OKAY;
_call(import_loc(locator, &lcp))
\end{verbatim}
\end {quote}

These are merely short-hand variables for accessing the data part of the LCP
(ie.\ everything except for the forward and backward links) and for accessing
the LCP state flags.

\begin {quote}
\begin {verbatim}
data               = &lcp->data;
state              = &data->state;
\end{verbatim}
\end {quote}

Ensure that there is no currently mapped data and that the object is primitive.

\begin {quote}
\begin {verbatim}
if (state->mapped)
        _call(DAT__PRMAP)
if (data->struc)
        _call(DAT__OBJIN)
\end{verbatim}
\end {quote}

Determine the shape of the object and match the dimensions. The {\tt
get\_shape} routine returns {\tt naxes} and the {\tt axis} array.

\begin {quote}
\begin {verbatim}
_call(get_shape(data, &naxes, axis))
if (*ndim != naxes)
        _call(DAT__DIMIN)
for (i=0; i<naxes; i++)
        if (dims[i] != axis[i])
                _call(DAT__DIMIN)
\end{verbatim}
\end {quote}

Validate the application data type specification. The {\tt check\_type} routine
copies the type from the supplied character variable, checking that there are
no illegal characters, ignoring white space and converting to upper-case.

\begin {quote}
\begin {verbatim}
_call(check_type(type, typbuf))
\end{verbatim}
\end {quote}

Determine the attributes of the application data and reject the operation if
not primitive. The {\tt get\_attrib} routine fills in the class, type and length
fields in the application data descriptor. The application must be requesting
data in one of the primitive types.

\begin {quote}
\begin {verbatim}
_call(get_attrib(typbuf, &data->app))
app                = &data->app;
if (app->dsc$b_class != DAT_K_PRIMITIVE)
        _call(DAT__TYPIN)
\end{verbatim}
\end {quote}

Match the object and application data attributes and reject the operation if
the types are incompatible.

\begin {quote}
\begin {verbatim}
obj                = &data->obj;
_call(match_types(obj, app))
\end{verbatim}
\end {quote}

Ensure that the object data is active. The {\tt rec\_\$peek} routine reads the
RCL from the object record.

\begin {quote}
\begin {verbatim}
_call(rec_$peek(&data->han, &rcl))
if (!rcl.active)
        _call(DAT__UNSET)
\end{verbatim}
\end {quote}

Save the address of the application program buffer, or use the descriptor
supplied by the program if character-type. The application data descriptor is
now complete.

\begin {quote}
\begin {verbatim}
app->dsc$a_pointer = values;
if (app->dsc$b_dtype == DSC$K_DTYPE_T)
        _chmove(8, values, app);
\end{verbatim}
\end {quote}

Calculate the length (in bytes) of the object data and determine the byte
offset into the object record's dynamic domain. Note that the object data
descriptor describes a single element of the object data.

\begin {quote}
\begin {verbatim}
objlen             = obj->dsc$w_length * data->size;
objoff             = obj->dsc$w_length * data->offset;
\end{verbatim}
\end {quote}

Gather the object data if discontiguous. The LCP may refer to a sliced object.
If so, the object data may be discontiguous. If so, the {\tt gather\_data}
routine will gather the discontiguous bits into a contiguous part of virtual
memory. The use of the {\tt copy\_status} variable is explained below.

\begin {quote}
\begin {verbatim}
if (state->broken)
        {
        gather_data(data);
        copy_status     = hds_gl_status;
        }
\end{verbatim}
\end {quote}

Otherwise, locate the record's dynamic domain and translate the data to the
program's buffer. The {\tt rec\_\$locate} and {\tt rec\_\$ditch} routines
always go in pairs. The {\tt rec\_\$locate} routine returns a pointer to a
contiguous part of the object record's dynamic domain. That pointer is
guaranteed to be valid until the {\tt rec\_\$ditch} routine is called.

\begin {quote}
\begin {verbatim}
else
        {
        _call(rec_$locate(&data->han, objlen, objoff, `R', &dom))
        obj->dsc$a_pointer = dom;
\end{verbatim}
\end {quote}

Both the object and application data descriptors are now complete. The {\tt
trans\_data} routine will translate the data from the object to the application
buffer. If no conversion is necessary then the data will simply be moved using
the {\tt OTS\$MOVE3} routine.

\begin {quote}
\begin {verbatim}
        trans_data(data->size, obj, app);
\end{verbatim}
\end {quote}

A conversion error may have set {\tt hds\_gl\_status}. However this is no
reason not to call {\tt rec\_\$ditch} so save {\tt hds\_gl\_status} in {\tt
copy\_status}, reset {\tt hds\_gl\_status}, call {\tt rec\_\$ditch}, and then
\ldots

\begin {quote}
\begin {verbatim}
        copy_status     = hds_gl_status;
        hds_gl_status   = DAT__OKAY;
        rec_$ditch(&data->han, objlen, objoff, `R', &dom);
        }
\end{verbatim}
\end {quote}
\ldots check the value of {\tt copy\_status} before return. If {\tt
copy\_status} is bad this has the effect of copying it to the status argument
and then exiting.

\begin {quote}
\begin {verbatim}
_call(copy_status)
\end{verbatim}
\end {quote}

This is the normal exit, returning the value of {\tt hds\_gl\_status} to the
caller.

\begin {quote}
\begin {verbatim}
_exit0;
}
\end{verbatim}
\end {quote}

\section {TUNING}
\label {system_tuning}

The {\tt HDS\_TUNE} routine is used to alter any of the six internal control
parameters and these will affect the behaviour of the system. The parameters
are discussed below.

\subsection {MAXWPL --- Maximum Size of the Working Page List}

Initially this is defaulted to 32 which provides 16 Kbytes of cache store. To
increase it --- it cannot be decreased --- {\tt HDS\_TUNE} must be called {\em
before} the system is started, eg:

\begin {quote}
\begin {verbatim}
.....startup HDS after allocating 100 pages for the WPL
CALL HDS_TUNE(`MAXWPL',100,STATUS)
CALL HDS_START(STATUS)
\end{verbatim}
\end {quote}

\subsection {INALQ --- Initial File Allocation Quantity}

This determines how many disc blocks are to be allocated when a new container
file is created. The default of 2 provides space for the Header Control Block
and one Logical Record Block (although, because of clustering, more than 2
blocks may actually be allocated). It must be set before each file creation as
it reverts to the default value afterwards, eg:

\begin {quote}
\begin {verbatim}
.....create file with initial size of 1030 blocks
CALL HDS_TUNE(`INALQ',1030,STATUS)
CALL HDS_NEW(`NGC1365',NAME,`plate',0,0,LOC,STATUS)
\end{verbatim}
\end {quote}

\subsection {NBLOCKS --- Transfer Size for File to File Copy}

This determines how many disc blocks are to be transferred per I/O operation
when performing a file to file copy. The default is 32 --- this can be
increased but cannot be decreased. It must be set before each file copy as it
reverts to the default value afterwards, eg:

\begin {quote}
\begin {verbatim}
.....copy file to a new container file
CALL HDS_TUNE(`NBLOCKS',100,STATUS)
CALL HDS_OPEN(`NGC1365',`READ',LOC,STATUS)
CALL HDS_COPY(LOC,`NGC1365',`NGC1365',STATUS)
\end{verbatim}
\end {quote}

\subsection {WAIT/SYSLCK --- Control over File-locking Facility}

{\bf WAIT} is a logical switch which specifies the action a process is to take
when attempting to lock for write-access a container file that is currently
locked by another process. It is initially set to FALSE, ie.\ the operation
returns immediately with an error status if the request cannot be granted. An
explicit call to {\tt HDS\_TUNE} must be made to alter this setting --- it does
not automatically revert to the default after being used.

{\bf SYSLCK} is a logical switch which, if set, allows container files to be
locked on a system-wide basis. The default of FALSE provides locking only on a
UIC group basis. Only processes with the VMS SYSLCK privilege can set this
parameter. See the section {\em Lock Management Services} in
\cite{system_services_reference} for more information on this subject.

\subsection {NCOMP --- Optimum Number of Components in a Structure}

This is used to specify a suitable value for the number of components in a new
structure (defaulted to 6), eg:

\begin {quote}
\begin {verbatim}
.....create and locate the RA structure object
CALL DAT_NEW(LOC1,`RA',`hms',0,0,STATUS)
CALL DAT_FIND(LOC1,`RA',LOC2,STATUS)

.....reset NCOMP to 3 before creating the first component
CALL HDS_TUNE(`NCOMP',3,STATUS)
CALL DAT_NEW(LOC2,`HOURS',`_byte',0,0,STATUS)
\end{verbatim}
\end {quote}

It reverts to the default value after the creation of a component record. It
reserves space in a newly created component record for subsequent entries.
Obviously, if a structure such as RA will always possess 3 components, a saving
of space can be made by decreasing {\tt NCOMP}. On the other hand, if a
structure with a large number of components is to be created, it may be
advisable to set this parameter appropriately to keep the component record
unchained for as long as possible. If not, the expansion of the dynamic domain
when the 7th object is created may force the contents into a PDB, which
prevents the domain from being cached.

\appendix

\clearpage
\section {Lists of Routines by Module}
\label {routine_lists}

In this section are listed all the routines in all the modules that comprise
HDS. They are split up into logical groups and the routines are listed
alphabetically within these groups. The main reason for providing these lists
is to aid in reading code.

For each routine, the argument list and a very brief description of its
function is given.

Note that no mention is made here of the PRO module. This is a set of routines
that can count the number of times HDS routines have been called and maintain
totals of the elapsed and CPU times spent in them. These routines are not
regarded as being part of HDS.

\subsection {High-level DAT Routines}

These routines split into the {\em environment} routines and the {\em
application} routines. None of the environment routines need normally be called
from applications since access to them is hidden behind calls to {\em get
parameter} and {\em put parameter} routines. They are listed in
Table~\ref{high_level_dat_routines_hds_routines} and the application routines
are listed in Table~\ref{high_level_dat_routines_dat_routines}.

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|l||}
\hline
Routine name and argument list		& Description \\
\hline
{\tt HDS\_CLOSE (locator,status)}	& Close container file \\
{\tt HDS\_COPY (locator,file,name,status)} & Copy object to container file \\
{\tt HDS\_ERASE (locator,status)}	& Erase container file \\
{\tt HDS\_FLUSH (group,status)}	& Flush locator group \\
{\tt HDS\_FREE (locator,status)}	& Free container file \\
{\tt HDS\_GROUP (locator,group,status)}	& Locator group \\
{\tt HDS\_LINK (locator,group,status)}	& Link locator to group\\
{\tt HDS\_LOCK (locator,status)}	& Lock container file \\
{\tt HDS\_NEW (file,name,type,ndim,dims,locator,status)} & Create new container file \\
{\tt HDS\_OPEN (file,mode,locator,status)} & Open existing container file \\
{\tt HDS\_PROFL (name,level,status)}	& Enable execution profiling \\
{\tt HDS\_RUN (subroutine,status)}	& Run application subroutine \\
{\tt HDS\_SHOW (topic,status)}		& Show HDS statistic \\
{\tt HDS\_START (status)}		& Startup locator facility \\
{\tt HDS\_STATE (state,status)}		& HDS state ? \\
{\tt HDS\_STOP (status)}		& Rundown locator facility \\
{\tt HDS\_TRACE (locator,nlev,path,file,status)} & Trace path of object \\
{\tt HDS\_TUNE (cont,value,status)}	& Tune HDS control setting \\
\hline
\end {tabular}
\caption {High-level DAT Routines --- HDS\_ routines}
\label {high_level_dat_routines_hds_routines}
\end {center}
\end {table}

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|l||}
\hline
Routine name and argument list				& Description \\
\hline
{\tt DAT\_ALTER (locator,ndim,dims,status)}		& Alter size of object \\
{\tt DAT\_ANNUL (locator,status)}			& Annul locator \\
{\tt DAT\_BASIC (locator,mode,pntr,length,status)}	& Map data (in basic machine units) \\
{\tt DAT\_CELL (locator1,nsub,subs,locator2,status)}	& Locate cell \\
{\tt DAT\_CLONE (locator1,locator2,status)}		& Clone locator \\
{\tt DAT\_COERC (locator1,ndim,locator2,status)}	& Coerce object shape \\
{\tt DAT\_CONV (locator,type,conv,status)}		& Type conversion possible ? \\
{\tt DAT\_COPY (locator1,locator2,name,status)}		& Copy object \\
{\tt DAT\_ERASE (locator,name,status)}			& Erase object \\
{\tt DAT\_FIND (locator1,name,locator2,status)}		& Find named component \\
{\tt DAT\_GET (locator,type,ndim,dims,values,status)}	& Read data \\
{\tt DAT\_GETx (locator,ndim,dims,values,status)}	& Read I/R/D/L/C data \\
{\tt DAT\_INDEX (locator1,index,locator2,status)}	& Index into component list \\
{\tt DAT\_LEN (locator,len,status)}			& Primitive length ? \\
{\tt DAT\_MAP (locator,type,mode,ndim,dims,pntr,status)}& Map data \\
{\tt DAT\_MAPx (locator,mode,ndim,dims,pntr,status)}	& Map I/R/D/L/C data \\
{\tt DAT\_MOULD (locator,ndim,dims,status)}		& Alter shape of object \\
{\tt DAT\_MOVE (locator1,locator2,name,status)}		& Move object \\
{\tt DAT\_NAME (locator,name,status)}			& Object name ? \\
{\tt DAT\_NCOMP (locator,ncomp,status)}			& Number of components ? \\
{\tt DAT\_NEW (locator,name,type,ndim,dims,status)}	& Create new component \\
{\tt DAT\_PAREN (locator1,locator2,status)}		& Object parent ? \\
{\tt DAT\_PRIM (locator,prim,status)}			& Primitive object ? \\
{\tt DAT\_PUT (locator,type,ndim,dims,values,status)}	& Write data \\
{\tt DAT\_PUTx (locator,ndim,dims,values,status)}	& Write I/R/D/L/C data \\
{\tt DAT\_RENAM (locator,name,status)}			& Rename object \\
{\tt DAT\_RESET (locator,status)}			& Reset object state \\
{\tt DAT\_RETYP (locator,type,status)}			& Retype object \\
{\tt DAT\_SHAPE (locator,maxdim,dims,actdim,status)}	& Object shape ? \\
{\tt DAT\_SIZE (locator,size,status)}			& Object size ? \\
{\tt DAT\_SLICE (locator1,nsub,lower,upper,locator2,status)} & Locate slice \\
{\tt DAT\_STATE (locator,state,status)}			& Object state ? \\
{\tt DAT\_STRUC (locator,struc,status)}			& Structure object ? \\
{\tt DAT\_TEMP (type,ndim,dims,locator,status)}		& Create temporary object \\
{\tt DAT\_THERE (locator,name,there,status)}		& Object there ? \\
{\tt DAT\_TYPE (locator,type,status)}			& Object type ? \\
{\tt DAT\_UNMAP (locator,status)}			& Unmap data \\
{\tt DAT\_VALID (locator,valid,status)}			& Valid locator ? \\
{\tt DAT\_VEC (locator1,locator2,status)}		& Vectorise object \\
\hline
\end {tabular}
\caption {High-level DAT Routines --- DAT\_ routines}
\label {high_level_dat_routines_dat_routines}
\end {center}
\end {table}

\subsection {Low-level DAT Routines}

These routines are not user-callable. They provide common services to the
high-level DAT routines and are listed in Table~\ref{low_level_dat_routines}.

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|l||}
\hline
Routine name and argument list			& Description \\
\hline
{\tt CHECK\_GROUP (group,buf)}			& Check group  \\
{\tt CHECK\_MODE (mode,buf)}			& Check mode  \\
{\tt CHECK\_NAME (name,buf)}			& Check name  \\
{\tt CHECK\_SHAPE (ndim,dims,odl)}		& Check shape information \\
{\tt CHECK\_TYPE (type,buf)}			& Check type  \\
{\tt COPY\_OBJECT (ncomp,src,src\_crv,des,des\_crv)} & Recursively copy an object \\
{\tt DEFUSE\_LCP (pntr)}	       		& Deactivate LCP  \\
{\tt ENCODE\_SUBS (nlim,nsub,subs,buf,nchar)}	& Encode subscript expression \\
{\tt ERASE\_OBJECT  (ncomp,src,src\_crv,des,des\_crv)} & Recursively erase an object \\
{\tt EXPORT\_LOC (locator,pntr)}		& Export a locator \\
{\tt FLUSH\_DATA (data)}			& Flush mapped data  \\
{\tt GATHER\_DATA (data)}			& Gather discontiguous data \\
{\tt GET\_ATTRIB (typbuf,att)}			& Determine attributes of object \\
{\tt GET\_OFF (ndim,dims,subs,offset)}		& Get offset (from subscripts) \\
{\tt GET\_SHAPE (data,naxes,axis)}		& Enquire object shape \\
{\tt GET\_SUBS (ndim,dims,offset,subs)}		& Get subscripts (from offset) \\
{\tt IMPORT\_LOC (locator,pntr)}		& Import a locator \\
{\tt MAKE\_SCRATCH ()}				& Create scratch container file \\
{\tt MATCH\_TYPES (obj,app)}			& Match data types \\
{\tt MOVE\_DATA (nval,imp,exp)}			& Move data \\
{\tt MOVE\_OBJECT (ncomp,src,src\_crv,des,des\_crv)} & Recursively move an object \\
{\tt REFILL\_FLQ ()}				& Refill Free Locator Queue  \\
{\tt SCATTER\_DATA (data)}			& Scatter discontiguous data \\
{\tt TRANS\_CHAR (nval,imp,exp)}		& Translate data to or from character format \\
{\tt TRANS\_DATA (nval,imp,exp)}		& Translate data \\
\hline
\end {tabular}
\caption {Low-level DAT Routines}
\label {low_level_dat_routines}
\end {center}
\end {table}

\clearpage
\subsection {High-level REC Routines}

These are the externally-visible REC routines. They are called by the high-
and low-level DAT routines and are listed in
Table~\ref{high_level_rec_routines}.

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|l||}
\hline
Routine name and argument list	& Description \\
\hline
{\tt REC\_\$ADOPT (han,par)}	& Adopt new parent \\
{\tt REC\_\$ATTACH (file,state,mode,rcl,han)} & Attach container file \\
{\tt REC\_\$CLEAR (han)}	& Clear domain(D) \\
{\tt REC\_\$CREATE (par,rcl,han)} & Create new record \\
{\tt REC\_\$DEMAND (han,fns)}	& Demand container filename \\
{\tt REC\_\$DETACH (han,disp)}	& Detach container file \\
{\tt REC\_\$DITCH (han,length,offset,mode,pntr)} & Ditch domain(D) \\
{\tt REC\_\$EXPAND (han,extent)} & Expand domain(D) \\
{\tt REC\_\$FCOPY (src,des)}	& Physical file copy \\
{\tt REC\_\$FREEVM (size,pntr)} & Free virtual memory  \\
{\tt REC\_\$GRABVM (size,pntr)} & Grab virtual memory  \\
{\tt REC\_\$IDENT (han,rid)}	& Identify record \\
{\tt REC\_\$LIST ()}		& List open-file statistics \\
{\tt REC\_\$LOCATE (han,length,offset,mode,pntr)} & Locate domain(D) \\
{\tt REC\_\$LOCK (han)}		& Lock container file \\
{\tt REC\_\$MATCH (han1,han2)}	& Match handles \\
{\tt REC\_\$PEEK (han,rcl)}	& Peek domain(C) \\
{\tt REC\_\$PROFL (name,level)}	& Enable execution profiling \\
{\tt REC\_\$READ (han,buf)}	& Read domain(S) \\
{\tt REC\_\$RUBOUT (han)}	& Rubout existing record \\
{\tt REC\_\$SHRINK (han,extent)}& Shrink domain(D) \\
{\tt REC\_\$START ()}		& Startup record facility \\
{\tt REC\_\$STICK (rid,kin,han)} & Stick handle on record \\
{\tt REC\_\$STOP ()}		& Rundown record facility \\
{\tt REC\_\$UNLOCK (han)}	& Unlock container file \\
{\tt REC\_\$WRITE (han,buf)}	& Write domain(S) \\
\hline
\end {tabular}
\caption {High-level REC Routines}
\label {high_level_rec_routines}
\end {center}
\end {table}

\subsection {Low-level REC Routines}

These routines provide common services to the
high-level REC routines and are listed in Table~\ref{low_level_rec_routines}.

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|l||}
\hline
Routine name and argument list		& Description \\
\hline
{\tt CLEAR\_BITMAP (bitmap,length,nbits,sbpos)} & Clear bit map \\
{\tt CLOSE\_FILE (iochan)}		& Close file \\
{\tt CREATE\_FILE (file,size,alq)}	& Create file \\
{\tt DELETE\_FILE (fid)}		& Delete file \\
{\tt DROP\_BLOC (slot,bloc)}		& Drop (release) Bloc  \\
{\tt EXTEND\_FILE (fid,size,alq)}	& Extend file \\
{\tt FIND\_FILE (file,fid,fns)}		& Find file \\
{\tt FIND\_SLOT (fid)}			& Find slot \\
{\tt FLUSH\_BLOC (bcp)}			& Flush (write-back) Bloc \\
{\tt FREE\_FRAME (slot,size,bloc)}	& Free frame \\
{\tt FREE\_SPACE (size,range)}		& Free space \\
{\tt GRAB\_FRAME (slot,size,bloc)}	& Grab frame \\
{\tt GRAB\_SPACE (size,range)}		& Grab space \\
{\tt GROW\_FRAME (slot,size,extra,bloc)} & Grow frame \\
{\tt LOCK\_SLOT (slot)}			& Lock slot \\
{\tt MAP\_FRAME (slot,size,bloc,mode,pntr)} & Map frame \\
{\tt OPEN\_FILE (fid,mode,iochan)}	& Open file \\
{\tt SCAN\_BITMAP (bitmap,length,nbits,sbpos)} & Scan bit map \\
{\tt SEEK\_BLOC (slot,bloc,mode,pntr)} & Seek (locate) Bloc \\
{\tt SEEK\_BLOC\_1 (slot,bloc,mode,pntr)} & Seek (locate) Bloc (latter part) \\
{\tt SET\_BITMAP (bitmap,length,nbits,sbpos)} & Set bit map \\
{\tt SHUT\_SLOT (slot)}			& Shut slot \\
{\tt UNLOCK\_SLOT (slot)}		& Unlock slot \\
{\tt UNMAP\_FRAME (slot,size,pntr)}	& Unmap frame \\
{\tt UPDATE\_FREE (slot,bloc,cbm)}	& Update Free (record stack) \\
\hline
\end {tabular}
\caption {Low-level REC Routines}
\label {low_level_rec_routines}
\end {center}
\end {table}

\clearpage
\subsection {EXC Routines}

These routines are all user-callable although normally {\tt EXC\_\$HDS}
and {\tt EXC\_\$VMS} would only be called by HDS. They provide control over
the handling and reporting of error conditions and are listed in
Table~\ref{exc_routines}.

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|l||}
\hline
Routine name and argument list		& Description \\
\hline
{\tt EXC\_\$FLAGS (flags,oldflags)}	& Set error reporting flags \\
{\tt EXC\_\$HANDL (handler,oldhandler)}	& Define error handler \\
{\tt EXC\_\$HDS (status)}		& HDS exception handler \\
{\tt EXC\_\$LEVEL (level,oldlevel)}	& Set threshold severity level \\
{\tt EXC\_\$LOG (msg,status)}		& Log exception \\
{\tt EXC\_\$MSG (status,len,msg)}	& Translate condition value \\
{\tt EXC\_\$PROFL (name,level)}		& Enable execution profiling \\
{\tt EXC\_\$VMS (status)}		& VMS exception handler \\
\hline
\end {tabular}
\caption {EXC Routines}
\label {exc_routines}
\end {center}
\end {table}

\subsection {QUE Routines}

These routines hardly deserve a mention. They provide the same services
as the VAX {\tt INSQUE} and {\tt REMQUE} instructions and are listed in
Table~\ref{que_routines}.

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|l||}
\hline
Routine name and argument list	& Description \\
\hline
{\tt INSQUE (entryy,pred)}	& Insert entry in queue \\
{\tt REMQUE (entryy,addr)}	& Remove entry from queue \\
\hline
\end {tabular}
\caption {QUE Routines}
\label {que_routines}
\end {center}
\end {table}

\clearpage
\section {Error Codes}

This section is modified from the equivalent section in \cite{sun_92}.

A positive status value is returned by any HDS routine which detects an error
condition. If a program is to perform specific tests on status values, the
HDS-supplied symbolic identifiers described below should be used rather than
explicit numerical constants. Following DEC conventions, symbolic status
identifiers are prefixed with the facility name and a double underscore. The
appropriate symbol definitions are contained in a file with the logical name
{\tt dat\_err}. The symbols defined in this file are shown and described in
Table~\ref{error_codes}.

\begin {table}[htbp]
\begin {center}
\begin {tabular}{||l|l||}
\hline
{\em Identifier}	& {\em Description}\\
\hline
{\tt DAT\_\_ACTIV}	& Already active\\
{\tt DAT\_\_ACCON}	& Access conflict\\
{\tt DAT\_\_COMEX}	& Component already exists\\
{\tt DAT\_\_CONER}	& Conversion error\\
{\tt DAT\_\_CONIN}	& Conversion invalid\\
{\tt DAT\_\_DELIN}	& Deletion invalid\\
{\tt DAT\_\_DIMIN}	& Dimensions invalid\\
{\tt DAT\_\_FATAL}	& Fatal internal error\\
{\tt DAT\_\_FILCK}	& File locked\\
{\tt DAT\_\_FILIN}	& File invalid\\
{\tt DAT\_\_FILNF}	& File not found\\
{\tt DAT\_\_FILPR}	& File protected\\
{\tt DAT\_\_GRPIN}	& Group invalid\\
{\tt DAT\_\_INCHK}	& Integrity check\\
{\tt DAT\_\_LOCIN}	& Locator invalid\\
{\tt DAT\_\_MODIN}	& Mode invalid\\
{\tt DAT\_\_NAMIN}	& Name invalid\\
{\tt DAT\_\_OBJNF}	& Object not found\\
{\tt DAT\_\_OBJIN}	& Object invalid\\
{\tt DAT\_\_PRMAP}	& Primitive data mapped\\
{\tt DAT\_\_RELIN}	& Relationship invalid\\
{\tt DAT\_\_SUBIN}	& Subscripts invalid\\
{\tt DAT\_\_TRUNC}	& Text truncated\\
{\tt DAT\_\_TYPIN}	& Type invalid\\
{\tt DAT\_\_UNSET}	& Primitive data undefined\\
{\tt DAT\_\_VERMM}	& Version mismatch\\
{\tt DAT\_\_WEIRD}	& Unknown error\\
\hline
\end {tabular}
\caption {Error Codes}
\label {error_codes}
\end {center}
\end {table}

The meaning of the errors associated with these codes is described in more
detail below:

\begin{description}
\item [ACTIV]:
Routine {\tt HDS\_START} has been called when HDS was already active.
\item [ACCON]:
A write operation has been rejected because the container file was originally
opened for read-only access.
\item [COMEX]:
An attempt to create a new component of a structure has failed because a
component of the same name already exists.
\item [CONER]:
One of the following conditions occurred during type conversion:
\begin{itemize}
\item Floating point overflow (when converting from double precision to single
precision).
\item Integer overflow.
\end{itemize}
\item [CONIN]:
Conversion is not allowed between the two data types specified.
\item [DELIN]:
The object to be deleted was a non-empty structure.
\item [DIMIN]:
One of the following conditions occurred:
\begin{itemize}
\item When creating an object, the number of dimensions specified exceeds the
system-imposed limit (currently 7), or one or more of the dimension sizes is
zero or negative.
\item When slicing an object, the number of dimensions specified exceeds the
system-imposed limit (currently 3), or an incorrect number of lower and upper
bounds has been specified.
\item An attempt to read, write, or map primitive data has failed because the
dimensions specified do not match the actual shape of the object.
\end{itemize}
\item [FATAL]:
An unrecoverable error has been detected within one of the HDS modules.
Supplementary information is normally associated with this condition which gives
some indication of the nature of the error.
\item [FILCK]:
An attempt to open a container file in WRITE or UPDATE mode has failed
because the file is currently open for write-access by another process.
\item [FILIN]:
The specified file does not conform to the structure of an HDS container file.
\item [FILNF]:
The specified file cannot be found, or the device, directory or filename does
not obey the syntax rules for a VMS file specification.
\item [FILPR]:
The file is protected from being accessed in the specified mode (or from being
deleted).
\item [GRPIN]:
The supplied text does not conform to the syntax of a group specification.
\item [INCHK]:
Typically, an attempt to access an object has failed because the object has been
deleted or moved through another locator, either in the same program or in
another process.
\item [LOCIN]:
One of the following conditions occurred:
\begin{itemize}
\item A non-CHARACTER variable has been specified as a locator (or a CHARACTER
variable of the wrong length).
\item The specified locator is not currently active, ie.\ not associated with an
object.
\item The specified locator had been generated by an HDS routine which did not
run to successful completion.
\end{itemize}
\item [MODIN]:
The supplied text is not a valid access mode.
\item [NAMIN]:
The supplied text does not conform to the syntax of a name specification.
\item [OBJNF]:
A request to locate a component of a structure, by name or position, has failed
because the object does not exist.
\item [OBJIN]:
An input locator points to an object which is not suitable for the type of
operation requested.
Specifically, when the locator is associated with one of the following:
\begin{itemize}
\item A structure object where a primitive is required (and vice versa).
\item An n-D object where a scalar (or cell) is required.
\item A slice or cell of an n-D object where the complete array is required.
\item A top-level object where a component object is required (and vice versa).
\end{itemize}
\item [PRMAP]:
A GET, PUT or MAP operation has been rejected because primitive data is
currently mapped to the specified locator.
\item [RELIN]:
An attempt to move an object has failed because the recipient structure is
contained in a different file.
\item [SUBIN]:
An attempt to locate a slice or cell of an n-D object has failed
because the subscript information references an element which is outside the
array bounds of the object (or a lower bound exceeds an upper bound).
\item [TRUNC]:
Reading or writing character data has resulted in the truncation of significant
(non-blank) characters.
\item [TYPIN]:
The supplied text does not conform to the syntax of a type specification.
\item [UNSET]:
An attempt to read primitive object data has failed because the values are
currently undefined.
\item [VERMM]:
The HDS version number of a container file exceeds the current version number
of the HDS software.
Typically, this signifies that the file has previously been created by a new
version of HDS that cannot be handled by a program linked with an older version.
\item [WEIRD]:
If this condition ever occurs, panic.
\end{description}

\begin {thebibliography}{XXXXXXX}
\addcontentsline{toc}{section}{References}
\bibitem[FORTRAN]{programming_in_fortran} {\em Programming in VAX FORTRAN},
AA-D034D-TE, DEC
\bibitem[LSN/32]{lsn_32} LSN/32 (RAL), {\em HDS --- Hierarchical Data System: C
version}
\bibitem[MODPROC]{creating_modular_procedures} {\em Guide to Creating Modular
Procedures on VAX/VMS}, AA-FB84A-TE, DEC
\bibitem[RTL]{run_time_library_reference} {\em VAX/VMS Run-Time Library
Routines Reference Manual}, AA-QLZ55-GZ, DEC
\bibitem[SUN/92]{sun_92} SUN/92, {\em HDS --- Hierarchical Data System:
Programmer's guide}
\bibitem[SYSSERV]{system_services_reference} {\em VAX/VMS System Services
Reference Manual}, AA-Z501A-TE, DEC
\end {thebibliography}

\section* {Glossary}
\addcontentsline{toc}{section}{Glossary}
\markboth{GLOSSARY}{GLOSSARY}

\begin {description}
\item [AAO]: Anglo-Australian Observatory.
\item [Array]: A object that has an associated dimensionality. The opposite
of a scalar.
\item [BCP]: Block Control Packet (qv).
\item [BID]: Block-ID (qv).
\item [Bitmap]: An array of bits where each bit indicates the availability or
unavailability of some resource. HDS uses bitmaps for maintaining records
of virtual memory usage and of chip usage within LRBs.
\item [BLISS-32]: A high-level DEC-specific system programming language that
was used for the first version of HDS.
\item [Block]: A disc block. HDS regards the first block in a disc file as
being block 1.
\item [Block Control Packet]: A data structure (illustrated in
Figure~\ref{structure_of_a_block_control_packet}) that controls access to a
cached disc block.
\item [Block-ID]: A data structure (illustrated in
Figure~\ref{structure_of_a_block_id}) that describes a block and its associated
file.
\item [Broken]: Used of the data associated with an LCP. Such data is said
to be broken if it is not a contiguous part of the data held in the container
file (because it has been sliced).
\item [C]: A high-level, widely and fairly portable system programming language
that is used for the current version of HDS.
\item [Cell]: A single element of an array.
\item [Chained]: Used of a record's dynamic domain. A dynamic domain is said
to be chained if there is not enough room for it in the same LRB as the
rest of the record, forcing it to be held in a frame of PDBs. The opposite
of un-chained.
\item [Chips]: The unit of storage allocation in an LRB.
\item [Component]: An object at one level below a structure object. A component
can itself be either a structure or a primitive object.
\item [Component Record]: A record that contains a list of a structure object's
component names and associated RIDs.
\item [Component Record Vector]: A data structure that describes one element
of the dynamic domain of a Component Record.
\item [Container File]: An HDS file. The term is used to underline the fact
that the file itself is just an container for objects.
\item [Container Record]: The second record in a container file. It contains
the name and RID of the file's top-level object.
\item [Control Domain]: The fixed size (10 bytes) part of a record that
contains the RCL.
\item [CRV]: Component Record Vector (qv).
\item [DAT]: The top-level HDS module. It contains all of the user-callable
routines (apart from the EXC routines).
\item [Data Object]: Any HDS object. It can be a structure or a primitive
object.
\item [Dynamic Domain]: The dynamically alterable part of a record. It can
be expanded and contracted after a record has been created. In a structure
object it contains the RIDs of the child objects. In a primitive object it
contains actual data.
\item [EXC]: The exception-handling module. It contains routines to control
the handling and reporting of error conditions.
\item [FCV]: File Control Vector (qv).
\item [FID]: File-ID (qv).
\item [File]: Unless otherwise stated, this always means an HDS container file.
\item [File Control Vector]: A data structure (illustrated in
Figure~\ref{structure_of_a_file_control_vector_slot}) that contains information
about a container file currently being accessed by HDS.
\item [File-ID]: A data structure (illustrated in
Figure~\ref{structure_of_a_file_id}) that contains an encoded form of a
container file name.
\item [FORTRAN]: FORmula TRANslation. An early scientific programming language
still widely used in research and engineering environments. It is the most
common language used for HDS applications. The current standard version is
called Fortran-77 and VAX Fortran is a superset of this.
\item [FPL]: Free Page List (qv).
\item [Frame]: A set of contiguous PDBs.
\item [Free Page List]: Those pages in the data cache that are not currently
in use.
\item [Free Space Stack]: The latter part of a container file's HCB. It
contains details of free blocks and chips in the container file.
\item [HAN]: Handle (qv).
\item [Handle]: A data structure (illustrated in
Figure~\ref{structure_of_a_handle}) that contains the information necessary
to access a given chip in a given block in a given file.
\item [HCB]: Header Control Block (qv).
\item [HDS]: Hierarchical Data System. To find out more, read this document.
\item [Header Control Block]: The first block of a container file. It contains
various header items and the free space stack.
\item [LCP]: Locator Control Packet (qv).
\item [LOC]: Locator (qv).
\item [Locator]: A data structure (illustrated in
Figure~\ref{structure_of_a_locator}) that is passed to and from application
routines and which allows access to the associated LCP and thus to the
associated object.
\item [Locator Control Packet]: A data structure (illustrated in
Figure~\ref{structure_of_a_locator_control_packet}) that contains cached
information about an object and extra information about how that object is
currently being accessed.
\item [Logical Record Block]: A disc block that is split up into a chip bitmap
and $15 \times 34$ byte chips. This is in contrast to a PDB, which is pure
data.
\item [LRB]: Logical Record Block (qv).
\item [Mapped]: Used of object data. An object is said to be mapped if the
application has been given a pointer to its data. The term does not refer
to the method of access from the container file.
\item [Name]: Object name. A text string of not more than 15 characters
(padded to the right with spaces).
\item [Object]: Data Object (qv).
\item [Object Descriptor Label]: A data structure (illustrated in
Figure~\ref{structure_of_an_object_descriptor_label}) that describes an object's
type and shape. It comprises the static domain of a Structure or Primitive
Record.
\item [Object Record]: A Structure or Primitive Record (qv). Its static domain
is an ODL and its dynamic domain is either a set of RIDs or else primitive data.
\item [ODL]: Object Descriptor Label (qv).
\item [Parent]: Record Parent (qv).
\item [PDB]: Pure Data Block (qv).
\item [PLT]: Primitive Lookup Table (qv).
\item [Primitive]: Having actual associated data values. The opposite of
structure.
\item [Primitive Lookup Table]: A data structure that relates primitive types
to numbers of bits per pixel, VMS descriptor types etc.
\item [Primitive Record]: A record describing a primitive object. Its static
domain is an ODL and its dynamic domain is actual primitive data.
\item [PRO]: The profiling module. It contains routines to count numbers
of entries to routines and to maintain total elapsed and CPU times. It is
not regarded as being part of HDS.
\item [Pure Data Block]: A disc block which is all used for data. This is
in contrast to an LRB, which is split up into 15 chips.
\item [QUE]: The queue-handling module. It contains simulations of the VAX
{\tt INSQUE} and {\tt REMQUE} instructions.
\item [RAL]: Rutherford-Appleton Laboratory
\item [RCL]: Record Control Label (qv).
\item [REC]: The record-access module. It contains routines that are called
only by the routines in the DAT module.
\item [Record Classes]: There are four of these: Container Record, Component
Record, Structure Record and Primitive Record (qv).
\item [Record Control Label]: A data structure (illustrated in
Figure~\ref{structure_of_a_record_control_label}) that comprises the control
domain of all records. It is a fixed size structure that contains information
such as the lengths of the static and dynamic domains.
\item [Record Domains]: The three logical parts of a record. The control
domain is of a fixed size (10 bytes) and contains the RCL. The static domain is
of a fixed size (chosen at record creation time and cannot exceed 498 bytes).
The dynamic domain is of a variable size.
\item [Record Parent]: The record that owns the record in question.
\item [Record-ID]: A data structure (illustrated in
Figure~\ref{structure_of_a_record_id}) that describes a given chip in a given
record.
\item [RGO]: Royal Greenwich Observatory.
\item [RID]: Record-ID (qv).
\item [RMS]: Record Management Services. These are a part of the VMS Operating
System.
\item [RTL]: Run-Time Library. These are routines that are automatically
available to all programs running under VMS.
\item [Scalar]: A object that consists of a single item with no associated
dimensionality. The opposite of an array.
\item [Shape]: The dimensionality of an object. It can be described by a
number of axes and, if that number is greater than zero, one dimension size
for each axis.
\item [Slice]: A subset of an object. It is described by giving a range of
up to the first three dimensions. If the data in a slice is not contiguous
then it is said to be broken (qv).
\item [Slot]: A file number. It is used to refer to one of the files that
are currently opened by HDS --- the slot is the entry number in the FCV.
\item [SRV]: Structure Record Vector (qv).
\item [Stack]: (Free Space) Stack (qv).
\item [Static Domain]: The fixed size (chosen at record creation time and
cannot exceed 498 bytes) part of a record.
\item [STK]: (Free Space) Stack (qv).
\item [Structure]: Not having actual associated data values but instead having
zero or more components. The opposite of primitive.
\item [Structure Record]: A record describing a structure object. Its static
domain is an ODL and its dynamic domain is a list of its elements' RIDs.
\item [Structure Record Vector]: A data structure that describes one element
of the dynamic domain of a Structure Record.
\item [Type]: Object type. A text string of not more than 15 characters
(padded to the right with spaces). If it begins with an underscore the object
is primitive. Otherwise it is a structure.
\item [Un-chained]: Used of a record's dynamic domain. A dynamic domain is said
to be un-chained if its dynamic domain is held in the same LRB as the rest of
the record. The opposite of chained.
\item [VAX]: Virtual Address eXtended. The sole range of machines on which
HDS is currently implemented.
\item [Vectorised]: Used of the data associated with an LCP. Such data is said
to be vectorised if an explicit routine call has forced it to be regarded
as a vector.
\item [Virtual Memory Pool]: A data structure (illustrated in
Figure~\ref{structure_of_the_virtual_memory_pool}) that allows access to
the virtual memory that is managed by HDS.
\item [VMP]: Virtual Memory Pool (qv).
\item [VMS]: Virtual Memory System. VAX/VMS is the Operating System under
which HDS has been implemented.
\item [Working Page List]: Those pages in the data cache that are currently
in use.
\item [WPL]: Working Page List (qv).
\end {description}

\section* {Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}
\markboth{ACKNOWLEDGEMENTS}{ACKNOWLEDGEMENTS}
The original version of HDS was designed and implemented by Dave Pearce and
Anton Walter at RAL during 1982 and 1983. It was written in BLISS-32, a
powerful systems implementation language which produces highly optimised object
code. It has been translated into C by Mike Lawden at RAL and William Lupton at
RGO / AAO --- see \cite{lsn_32} for further details.

Documentation on the design and implementation of HDS was originally written
by Dave Pearce (RAL).
This was revised, reorganised and extended by Mike Lawden (RAL) and issued as
an earlier version of this paper.
\end {document}
