\documentstyle[11pt]{article} 
\pagestyle{myheadings}

%------------------------------------------------------------------------------
\newcommand{\stardoccategory}  {Starlink User Note}
\newcommand{\stardocinitials}  {SUN}
\newcommand{\stardocnumber}    {7.4}
\newcommand{\stardocauthors}   {S M Beard \\ P M Allan}
\newcommand{\stardocdate}      {31 March 1993}
\newcommand{\stardoctitle}     {GENERIC --- A Utility for Compiling Generic
				\\[2ex] Fortran Subroutines}
%------------------------------------------------------------------------------

\newcommand{\stardocname}{\stardocinitials /\stardocnumber}
\renewcommand{\_}{{\tt\char'137}}     % re-centres the underscore
\markright{\stardocname}
\setlength{\textwidth}{160mm}
\setlength{\textheight}{230mm}
\setlength{\topmargin}{-2mm}
\setlength{\oddsidemargin}{0mm}
\setlength{\evensidemargin}{0mm}
\setlength{\parindent}{0mm}
\setlength{\parskip}{\medskipamount}
\setlength{\unitlength}{1mm}

%------------------------------------------------------------------------------
% Add any \newcommand or \newenvironment commands here
%------------------------------------------------------------------------------

\begin{document}
\thispagestyle{empty}
SCIENCE \& ENGINEERING RESEARCH COUNCIL \hfill \stardocname\\
RUTHERFORD APPLETON LABORATORY\\
{\large\bf Starlink Project\\}
{\large\bf \stardoccategory\ \stardocnumber}
\begin{flushright}
\stardocauthors\\
\stardocdate
\end{flushright}
\vspace{-4mm}
\rule{\textwidth}{0.5mm}
\vspace{5mm}
\begin{center}
{\Large\bf \stardoctitle}
\end{center}
\vspace{5mm}

%------------------------------------------------------------------------------
%  Add this part if you want a table of contents
  \setlength{\parskip}{0mm}
  \tableofcontents
  \setlength{\parskip}{\medskipamount}
  \markright{\stardocname}
%------------------------------------------------------------------------------

\newpage
\section{Introduction}

GENERIC is a utility which preprocesses a generic Fortran subroutine into
its different types and concatenates these routines into a file.
The file can then be compiled with the Fortran compiler to produce an
object module.

\section{GENERIC Subroutines - an Example}

A generic subroutine is one which has an argument (or arguments)
which may be one of a number of different Fortran data types.

As an example, consider a subroutine which initialises every
element in an array to zero. 
One might want a different version of the routine for initializing
a double precision array, an integer array and a real array.
The following subroutines would need to be written:

\begin{verbatim}
      ZEROD( N, DARR )   Zero every element in a DOUBLE PRECISION array
      ZEROI( N, IARR )   Zero every element in an INTEGER array
      ZEROR( N, RARR )   Zero every element in a REAL array
\end{verbatim}

The integer routine might contain the following code (in a file
called {\tt zeroi.f} on Unix; ZEROI.FOR on VMS):

\begin{verbatim}
            SUBROUTINE ZEROI( N, ARRAY )
      *+
      *  Name:
      *     ZEROI
      *  Purpose:
      *     Zero all the elements of an INTEGER array
      *  Invocation :
      *     CALL ZEROI( N, ARRAY )
      *  Description :
      *     This sets all the elements of an INTEGER array to zero
      *  Arguments:
      *     N = INTEGER( Given )
      *        Dimension of array
      *     ARRAY( N ) = INTEGER( Given and Returned )
      *        The array to be zeroed
      *                   "
      *     (the rest of the standard prologue)
      *                   "
      *  Type Definitions:
            IMPLICIT NONE
      *  Arguments Given:
            INTEGER N
      *  Arguments Given and Returned:
            INTEGER ARRAY( N )
      *  Local Variables:
            INTEGER I                   ! Array index
      *-
      
            DO I = 1, N      
               ARRAY( I ) = 0
            END DO
            END
\end{verbatim}

The REAL and DOUBLE PRECISION versions of the routine would have to be
written in a similar way, and this obviously involves a lot of 
duplication of effort.
The situation gets even worse if other versions of the routine
(such as BYTE, INTEGER*2 etc...) are needed.
Also, it is very difficult and tedious to keep all the versions
in step if they have to be edited individually when the routines
are changed.

The GENERIC utility is a labour saving device which enables all the
various types of routines to be constructed automatically from
one master routine.
If any changes become necessary, only this one master routine needs
to be edited.

To use the GENERIC utility, the routine listed above is replaced by
the following (in a file called {\tt zero.gen} on Unix; ZERO.GEN on VMS):

\begin{verbatim}
            SUBROUTINE ZERO<T>( N, ARRAY )
      *+
      *  Name:
      *     ZERO<T>
      *  Purpose:
      *     Zero all the elements of a <COMM> array
      *  Invocation :
      *     CALL ZERO<T>( N, ARRAY )
      *  Description :
      *     This sets all the elements of a <COMM> array to zero
      *  Argument:
      *     N = INTEGER( Given )
      *        Dimension of array
      *     ARRAY( N ) = <TYPE>( Given and Returned )
      *        The array to be zeroed
      *                   "
      *     (the rest of the standard prologue)
      *                   "
      *  Type Definitions:
            IMPLICIT NONE
      *  Arguments Given:
            INTEGER N
      *  Arguments Given and Returned:
            <TYPE> ARRAY( N )
      *  Local Variables:
            INTEGER I                  ! Array index
      *-
      
            DO I = 1, N
               ARRAY( I ) = 0<CONST>
            END DO
            END
\end{verbatim}

This is a ``generic routine.''
Other examples of generic routines include the DAT\_ routines in
the Hierarchical Data System (SUN/92) and the PAR\_ routines of
the ADAM parameter system (SUN/94).

The items in angle brackets, $<>$ , are ``tokens'' which the generic
utility replaces when it converts the routine into one of a given
type.
In the above example, the tokens would be replaced
as follows :-

\begin{verbatim}
       Token       Double precision    Integer       Real
                   replacement         replacement   replacement

       <T>         D                   I             R
       <TYPE>      DOUBLE PRECISION    INTEGER       REAL
       <COMM>      DOUBLE PRECISION    INTEGER       REAL
       <CONST>     .0D0                (blank)       .0E0
\end{verbatim}

The types are generated by issuing the following command from DCL on a VMS
system :-

\begin{quote}{\tt
\$ GENERIC ZERO.GEN/TYPES=(D,I,R)
}
\end{quote}

or this command on a Unix system :-

\begin{quote}{\tt
\% generic -t dir zero.gen
}
\end{quote}

More details on the use of the GENERIC utility, and a full list of 
tokens and types, are given in the rest of this document.

\section{Features of the GENERIC Utility}

\subsection{The KERNEL Tokens}

The following tokens are recognised by the GENERIC utility :-

\newlength{\numlen}
\settowidth{\numlen}{000000000000}
\settowidth{\labelsep}{000}

\begin{list}{}{\setlength{\labelwidth}{\numlen}\setlength{\leftmargin}{\numlen}
\addtolength{\leftmargin}{\labelsep}}

\item[\verb+<T>+] To be replaced by a single character representation
of the specified type.

\item[\verb+<TYPE>+] To be replaced by the string used to declare variables
of the specified type when they are passed as arguments to
the subroutine.

\item[\verb+<LTYPE>+] To be replaced by the string used to declare variables
of the specified type when they are defined as local variables
in the subroutine.

\item[\verb+<CONST>+] To be replaced by the string appended to numeric constants
of the specified type.

\item[\verb+<HTYPE>+] To be replaced by the HDS name for the specified type.

\item[\verb+<COMM>+] To be replaced by a comment describing the specified type
in plain English.

\end{list}

A full table of the translation of these tokens
for each variable type is given in
Appendix 1 at the end of this document. 

\subsection{The ASTERIX Tokens}

As well as the kernel tokens, there are an additional set of
tokens which will be translated if the ASTERIX option is specified
with the GENERIC command (see below).

These tokens are available to allow GENERIC to process existing
ASTERIX software (SUN/98) which uses them.
They are {\em not}\ recommended for use by new software.
A list of the ASTERIX tokens is given in Appendix 1.

\subsection{The Local CHARACTER Size}

The \verb+<LTYPE>+ token is used for declaring a variable which is local
to the subroutine in which it is used, whereas \verb+<TYPE>+ is used
to declare a variable which is passed to the routine as an argument.
In practice, the translation for both of these tokens is identical
for all types except CHARACTER.

In the CHARACTER type, a variable passed as an argument is declared
CHARACTER*(*), but a local variable is declared CHARACTER*\verb+<CHASIZ>+.
\verb+<CHASIZ>+ is replaced by the ``local character size.''
If not told otherwise, GENERIC will replace \verb+<CHASIZ>+ by 200,
making local character variables CHARACTER*200.

It is possible to override this default character size
if it is unsuitable.
This can be done using the GENSETDEF command, or by specifying
a /CHASIZ qualifier to the GENERIC command on a VMS system (see below).
On a Unix system, you can define the environment variable GENERIC\_CHASIZ or
you can use the {\bf -c} flag on the {\bf generic} command.

\subsection{The Maximum Record Length}
 
The GENERIC utility assumes it is dealing with Fortran 77
source code which cannot go beyond column 72.
If the substitution of a token results in a line becoming longer
than 72 characters, it will be broken automatically at column 72
and continued on the next line, using a ``:'' continuation character
in column 6 of the next line.

The VMS version of GENERIC will issue a warning message if it has to do this.
Getting this message is not serious, but it warns that, should the
routine in question fail, the line numbers referred to in a traceback
will not correspond to line numbers in the original generic routine.

It is recommended that source lines in generic routines be kept
well below 72 characters in length to prevent this from happening.

{\bf NOTE:}~{\it Comment lines (~{\rm i.e.} those beginning `C', `c', `D', `d' 
or `*') which are longer than 72 characters are unaltered by GENERIC.
So are lines of code with an in-line comment (~{\rm i.e.} those containing
a ``!'' character).}

This algorithm is not perfect and it is possible to fool it with end of line
comments on long lines or exclamation marks in character strings. However, this
seems not to have proven a problem in practice.

\section{Using the GENERIC Utility on VMS}

\subsection{The GENERIC Command}

The GENERIC utility is executed by typing the following command:

\begin{quote}{\tt
\$ GENERIC  <file-spec>[/qualifier[/qualifier...]]
}
\end{quote}

\verb+<file-spec>+ is the name of the file containing the generic routine.
The default type for such a file is ``{\tt .GEN}.''
The various routines generated are written to a file of type ``{\tt .FOR}'',
which is then compiled into an object module of type ``{\tt .OBJ}.''
The {\tt .FOR} file is then deleted (though this may be prevented by using
the /NOCOMPILE qualifier, below).

Optional qualifiers may be appended to the file-spec.
These can be classified into two types :-

\paragraph{Compiler qualifiers :}
Any qualifier for the Fortran compiler may be given
(e.g. /LIST, /DEBUG, /D\_LINES etc...). (See the VMS Fortran manual,
or type HELP FORTRAN for more details on these).

\paragraph{Generic qualifiers :}
In addition, there are certain qualifiers which the
GENERIC utility itself will recognise. These are:

\begin{description}
\begin{description}
\item[/TYPES=(type-list)]
This specifies the types required. The types should be separated
by commas and enclosed in parentheses (e.g. /TYPES=(C,D,I,L,R) )
The following types are possible :-
\begin{center}
\begin{tabular}{ll}
	B  &- Byte \\
	UB &- Unsigned byte \\
	C  &- Character \\
	D  &- Double precision \\
	I  &- Integer \\
	L  &- Logical \\
	R  &- Real \\
	W  &- Word \\
	UW &- Unsigned word \\
\end{tabular}
\end{center}
There are also some key words which can be used in the /TYPES qualifier for 
convenience. These are :-
\begin{verbatim}
	/TYPES=ALL		! Generate all types
	/TYPES=NUMERIC		! Generate all numeric types
\end{verbatim}
If the /TYPES qualifier is not specified, then the list of types which
is currently defined as the default (using GENSETDEF) will be generated.
If a default list of types has not been specified then all types will be
generated. 

\item[/CHASIZ=local-character-size]
This specifies the local character size required.
This is the size (in bytes) of any CHARACTER variables which are
declared as local variables (with the \verb+<LTYPE>+ token) in a GENERIC
routine. 

If the /CHASIZ qualifier is not specified, then the local
character size which has been defined as the default (using GENSETDEF)
will be used.
If a default size has not been specified, then 200 will be used
(resulting in CHARACTER*200).

\item[/ASTERIX]
If specified, this qualifier causes GENERIC
to replace the ASTERIX tokens (listed in Appendix 1),
as well as the usual (KERNEL) tokens.
Using this qualifier will slow down the execution of GENERIC,
because it will need to search for more tokens.

\item[/NOCOMPILE]
If present, this qualifier specifies that the resultant
Fortran source file should be kept, and not compiled and deleted.

\end{description}
\end{description}

\subsection{The GENSETDEF Command}

The command GENSETDEF can be used to set up default parameter values
for GENERIC to use subsequently.
The format of this command is as follows :-

\begin{quote}{\tt
\$ GENSETDEF  <default-types>  <default-local-character-size>
}
\end{quote}

If any parameter is not given it will be prompted for,
and it is usually easier to let GENSETDEF prompt for what it needs.

\verb+<default-types>+ is a string of types, in the same format as given to the
/TYPES qualifier (but without the parentheses), which GENERIC is to use
subsequently whenever a /TYPES qualifier is not specified. 

\verb+<default-local-character-size>+ is a default local character size
to be used subsequently whenever the /CHASIZ qualifier is not
specified.
(This will override the 200 initially used as the default.)

All the defaults set up by GENSETDEF will remain for the login
session, but will disappear on logout.

\subsection{Using GENERIC\_PROG Directly}

The GENERIC utility {\tt GENERIC\_DIR:GENERIC.COM} uses a program called \\
{\tt GENERIC\_DIR:GENERIC\_PROG.EXE} to produce the Fortran code file.
For special purposes, this program may be executed on its own.
(The tables in Appendix 1 were generated using {\tt GENERIC\_PROG}).
It is executed as a DCL foreign command as follows :-

\begin{verbatim}
      $ GENERIC_PROG :== $GENERIC_DIR:GENERIC_PROG
      $ GENERIC_PROG <input> <output> <types>[/qualifier] <chasiz>
\end{verbatim}

\verb+<input>+ specifies the name of the input file.
	
\verb+<output>+ specifies the name of the output file.

\verb+<types>+ specifies the types to be generated (in the same form
as used in the /TYPES qualifier for GENERIC (but without the 
parentheses).
The /ASTERIX qualifier only may be appended to the types string.

\verb+<chasiz>+ specifies the local character size to be used.

If any of the first three parameters are not specified,
then the GENERIC\_PROG program will prompt for them.

\subsection{Using the GENERIC Utility from LIBMAINT}

LIBMAINT is a procedure for maintaining modules within a text/object
library pair, and is described in SUN/99.
The GENERIC utility is designed to integrate easily with LIBMAINT,
and it is treated by LIBMAINT just like another compiler.
Generic modules are recognised by having type ``{\tt .GEN}''
appended to their names.
Note that specifying ``{\tt .GEN}'' after a module name is compulsory
in this case in order to tell LIBMAINT that the module is
generic.

The types which GENERIC is to generate can be defined by using /TYPES
as a compiler qualifier in LIBMAINT (see SUN/99), or by using GENSETDEF.
Using GENERIC with LIBMAINT is best illustrated by some examples.
Suppose the library ABCLIB contains generic modules, such as
ABC\_GET0.GEN, ABC\_GET1.GEN, ABC\_PUT0.GEN, ABC\_PUT1.GEN etc...

\begin{enumerate}

\item To insert a new generic module, producing C,D,I,L \& R types
(note the double slash ``//'' and the compulsory ``{\tt .GEN}'') :-

\begin{quote}{\tt
\$ LIBMAINT ABCLIB \\
\$ INSERT ABC\_GET0.GEN//TYPES=(C,D,I,L,R)
}
\end{quote}


\item To recompile all the modules, generating C,D,I,L \& R types for
the generic ones :-

\begin{quote}{\tt
\$ LIBMAINT \\
\$ GENSETDEF "C,D,I,L,R" 200 \\
\$ OLBCRE ABCLIB
}
\end{quote}

\end{enumerate}
The OBJLIST command in LIBMAINT can be used to verify which types
have been generated.

\section{Using the GENERIC Utility on Unix}

The GENERIC utility is executed by typing the following command:

\begin{quote}{\tt
\% generic [-t <types>] [-c <chasiz>] [-s] [-sc <char>] [-a] file1 [file2...]
}
\end{quote}

Arguments in square brackets are optional and items in angle brackets are
descriptive terms for the actual string that would be given. The simplest form
of the {\tt generic} command is:

\begin{quote}{\tt
\% generic prog.gen
}
\end{quote}

which will produce a file called {\tt prog.f} containing routines for all of
the supported data types. Note that the Unix version of generic does {\em not}
compile the Fortran files. All generic input files must have a suffix of
{\tt .gen}. You can specify multiple input files explicitly or with wild cards.
All output files have a suffix of {\tt .f}.

The generic command accepts the following flags:

\begin{description}
\item[-t] This must be followed by a list of the data type to be processed. The
list should be separated from the {\tt -t} by a space, and should consist of
one or more of the letters a, b, B, c, d, i, l, n, r, w or W. The letters have
the following meanings:
\begin{center}
\begin{tabular}{ll}
a & all data types \\
b & BYTE \\
B & unsigned BYTE \\
c & CHARACTER \\
d & DOUBLE PRECISION \\
i & INTEGER \\
l & LOGICAL \\
n & all numeric types (i.e. everything except CHARACTER and LOGICAL)\\
r & REAL \\
w & WORD (i.e. INTEGER*2) \\
W & unsigned WORD \\
\end{tabular}
\end{center}

For example, to generate routines for all of the standard FORTRAN numeric data
types, type the command:

\begin{quote}{\tt
\% generic -t dir prog.gen
}
\end{quote}

(Using {\tt n} instead of {\tt dir} would generate byte and word versions as
well.)

\item[-c] This should be followed by a space and then an integer giving the
value of the local character size required. The default value is 200.

\item[-s] Generate the output in separate files for each data type with the
single character type specifier appended to the base name of the file. The
default is to generate one output file for each input file. For example, if the
input file is {\tt sub1.gen} and the {\tt -t} flag is set to {\tt dir}, then
specifying the {\tt -s} flag will cause output files called {\tt sub1d.f}, {\tt
sub1i.f} and {\tt sub1r.f} to be generated. This option is provided as it is
usual on Unix systems to have all Fortran routines in separate source files.

\item[-sc] This flag allows the user to specify one or more characters that
will be used to separate the base file name from the appended character when
generating separate output files for each data type. The {\tt -sc} should be
followed by a space and then the character(s) to be used as the separator. For
example, the command:

\begin{quote}{\tt
\% generic -t dir -sc \_ func2.gen
}
\end{quote}

will generate files called {\tt func2\_d.f}, {\tt func2\_i.f} and {\tt
func2\_r.f}. If the {\tt -sc} flag is given, then the {\tt -s} flag can be
omitted.

\item[-a] This will cause the ASTERIX tokens to be interpreted as well as the
standard ones. Its use is deprecated.

\end{description}

It is possible to specify the list of data types to be processed and the local
character size by setting environment variables {\tt GENERIC\_TYPES} and {\tt
GENERIC\_CHASIZ}, e.g.

\begin{quote}{\tt
\% setenv GENERIC\_TYPES cdilr \\
\% setenv GENERIC\_CHASIZ 300
}
\end{quote}

{\bf N.B.} {\em The Unix version of generic does {\bf NOT} compile the
resulting FORTRAN output files. This must be done separately.}


\section{Data Type Conversions -- the DCV Routines}

\subsection{Introduction}

Occasionally, it may be necessary to convert data passed into
a generic routine in a generic variable into another type.
A general set of conversion functions is provided to make this
possible.
A naming convention is used so that the function
DCV\_\verb+<T1>TO<T2>+ converts data type \verb+<T1>+ into \verb+<T2>+.

The functions which can be are defined using Fortran statement 
functions in a file which may be included in a program by;

\begin{quote}{\tt
INCLUDE 'DCV\_FUN'
}
\end{quote}

which should come after any data type definitions, but before
any DATA initialisation or executable code.
In the standard prologue of SGP/16, it should go in the
``Internal references'' section.

Conversions which cannot be defined as statement functions are provided as
external functions in an object library ({\tt GENERIC\_DIR:DCV.OLB on VMS;
/star/lib/dcv.a on Unix}). Programs
requiring these functions should be linked:

\begin{quote}{\tt
\$ LINK program,GENERIC\_DIR:DCV/LIB
}
\end{quote}

on VMS, or compiled with

\begin{quote}{\tt
\% f77 prog.f -L/star/lib `dcv\_link`
}
\end{quote}

on Unix.

The types of these functions can be declared in a program by including;

\begin{quote}{\tt
INCLUDE 'DCV\_EXT'
}
\end{quote}

(in the ``External references'' section of the standard prologue).

{\bf NOTE:}~{\it Statement functions are considerably faster than external 
functions, so programs which use the common conversions defined in `DCV\_FUN'
will execute a lot faster.}

Only numeric conversions are supported at present, and
there are no DCV\_ functions for conversions involving CHARACTER and
LOGICAL data types.
A full list of the functions, the conversion performed, and their
location and restrictions, is given in Appendix 2.

\subsection{An Example of Data Type Conversion}

As an example, a generic subroutine for finding the mean of the data
contained in an array, where the array is of varying type but the mean is
always REAL, might look like the following :- 

\begin{verbatim}
            SUBROUTINE MEAN<T>( N, ARRAY, MEAN )
      *+
      *  Name:
      *     MEAN<T>
      *  Purpose:
      *     Calculate the mean value contained in a <COMM> array.
      *  Invocation :
      *     CALL MEAN<T>( N, ARRAY, MEAN )
      *  Description :
      *     This calculates the mean value of then data contained in the given
      *     <COMM> array, and returns the value in the REAL variable MEAN.
      *  Arguments:
      *     N = INTEGER( Given )
      *        The dimension of the array.
      *     ARRAY( N ) = <TYPE>( Given )
      *        The array containing the data to be averaged.
      *     MEAN = REAL( Returned )
      *        The mean of the values contained in the array.
      *                   "
      *     (the rest of the standard prologue)
      *                   "
      *  Type Definitions:
            IMPLICIT NONE
      *  Arguments Given:
            INTEGER N
            <TYPE> ARRAY( N )
      *  Arguments Returned:
            REAL MEAN
      *  External References:
            INCLUDE 'DCV_EXT'          ! Data conversion external functions
      *  Local Variables:
            INTEGER I                  ! Array index
            <LTYPE> SUM                ! Sum of the data in the array
      *  Internal References:
            INCLUDE 'DCV_FUN'          ! Data conversion statement functions
      *-
      
      *  Initialise the sum to zero
            SUM = 0<CONST>
      
      *  Accumulate the sum of the contents of the array
            DO I = 1, N
              SUM = SUM + ARRAY( N )
            END DO
      
      *  Divide by the dimension of the array to obtain the mean, using
      *  DCV_<T>TOR to convert from <COMM> into REAL.
            MEAN = DCV_<T>TOR( SUM ) / REAL( N )

            END
\end{verbatim}

In the above example, for the INTEGER version of the routine for
instance, the DCV\_ITOR(~SUM~) statement will be converted to
REAL(~SUM~) by the statement function in `DCV\_FUN'. 

\subsection{Range Checking}

The DCV\_\verb+<T1>+TO\verb+<T2>+ functions will not check the range of the 
values given to them, and overflow errors are possible from some conversions
(e.g. INTEGER to BYTE).
It is recommended that an application check the range of any value
given to a DCV function, to ensure it is within the allowed range.
Global parameters of the form DCV\_\_\verb+<T>+MIN and DCV\_\_\verb+<T>+MAX 
are available for this purpose, which give the allowed range for type
\verb+<T>+ ({\it e.g.} -128 to +127 for a BYTE).
These by may included in a program using;

\begin{quote}{\tt
INCLUDE 'DCV\_PAR'
}
\end{quote}

(in the ``Global parameters'' section of the standard prologue).

An example of the use of these parameters (for instance
converting an INTEGER to another type could be) :-

\begin{verbatim}
      *             "
      *     (standard prologue)
      *             "
      *  Global parameters:
            INCLUDE 'DCV_PAR'          ! DCV global parameters giving allowed ranges
      *             "
      *  External References:
            INCLUDE 'DCV_EXT'          ! DCV external function definitions
      *             "
      *  Local Variables:
            INTEGER IVAL               ! Integer value to be converted
            <TYPE> VAL                 ! Destination variable of type <COMM>
                    "
      *  Internal References:
            INCLUDE 'DCV_FUN'          ! DCV in-line statement functions
      *-
                    "
                    "
            IF ( ( IVAL .GE. DCV__<T>MIN ) .AND.
           :     ( IVAL .LE. DCV__<T>MAX ) ) THEN

      *  Value within range. The conversion is allowed.
               VAL = DCV_ITO<T>( IVAL )

            ELSE
      
      *  Error. Value out of range ...
               <take appropriate action>
            END IF
                    "
                    "
\end{verbatim}
      
Note that ranges are only provided for numeric data types.

An alternative (and easier) way of preventing numerical overflow from causing
problems is to use the PRIMDAT routines instead of the DCV ones. The PRIMDAT
routines will not crash your program if an overflow occurs. See SUN/39 for
details.

\newpage
\begin{center}
{\Large\bf APPENDIX 1}

{\large\bf A FULL TABLE OF TOKEN TRANSLATIONS}
\end{center}

This appendix gives the translations of all the tokens
made by GENERIC for each of the data types.
If a translation is illegal, the result will be ``?''
(and GENERIC will give a warning message).

To produce the BYTE version of a program, the following substitutions
are made :-
%bytekernal
\begin{tabbing}
xxxxxx\=xxxxxxxxxxxxx\=xxx\=xxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxx\kill 
\>{\bf Kernel} tokens (translated by default) :- \\
\\
\>TOKEN  \>SUBSTITUTE  \>\>COMMENT \\
\\
\>$<$T$>$       \>\>{\tt "B"}       \>{\em Subroutine name extension} \\
\>$<$TYPE$>$    \>\>{\tt "BYTE"}    \>{\em Argument type definition} \\
\>$<$LTYPE$>$   \>\>{\tt "BYTE"}    \>{\em Local type definition} \\
\>$<$CONST$>$   \>\>{\tt " "}       \>{\em String to be appended to constant} \\
\>$<$HTYPE$>$   \>\>{\tt "\_BYTE"}  \>{\em HDS type} \\
\>$<$COMM$>$    \>\>{\tt "BYTE"}    \>{\em Comment describing type} \\
\\
\>{\bf ASTERIX} tokens (translated when /ASTERIX is specified) :- \\
%byteasterix
\\
\>TOKEN  \>SUBSTITUTE  \>\>COMMENT \\
\\ 
\>$<$CONV$>$    \>\>{\tt " "} \>{\em BYTE to REAL conversion} \\
\>$<$EXT$>$     \>\>{\tt " "} \>{\em BYTE to REAL conversion extension} \\
\>$<$BRAK$>$    \>\>{\tt " "} \>{\em Brackets for BYTE to REAL conversion} \\
\>$<$ICONV$>$   \>\>{\tt " "} \>{\em BYTE to INTEGER conversion} \\
\>$<$IEXT$>$    \>\>{\tt " "} \>{\em BYTE to INTEGER conversion extension} \\
\>$<$IBRAK$>$   \>\>{\tt " "} \>{\em Brackets for BYTE to INTEGER conversion} \\
\>$<$REV$>$     \>\>{\tt " "} \>{\em REAL to BYTE conversion} \\
\>$<$RVBRAK$>$  \>\>{\tt " "} \>{\em Brackets for REAL to BYTE conversion} \\
\>$<$IREV$>$    \>\>{\tt " "} \>{\em INTEGER to BYTE conversion} \\
\>$<$IRVBRAK$>$ \>\>{\tt " "} \>{\em Brackets for INTEGER to BYTE conversion} \\                          
\end{tabbing}
To produce the UNSIGNED BYTE version of a program, the following
substitutions are made :-
%usbytekernal
\begin{tabbing}
xxxxxx\=xxxxxxxxxxxxx\=xxx\=xxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxx\kill 
\>{\bf Kernel} tokens (translated by default) :- \\
\\ 
\>TOKEN     \>SUBSTITUTE     \>\>COMMENT \\
\\
\>$<$T$>$       \>\>{\tt "UB"}     \>{\em Subroutine name extension} \\
\>$<$TYPE$>$    \>\>{\tt "BYTE"}   \>{\em Argument type definition} \\
\>$<$LTYPE$>$   \>\>{\tt "BYTE"}   \>{\em Local type definition} \\
\>$<$CONST$>$   \>\>{\tt " "}      \>{\em String to be appended to constant}\\
\>$<$HTYPE$>$   \>\>{\tt "\_UBYTE"}  \>{\em HDS type} \\
\>$<$COMM$>$    \>\>{\tt "UNSIGNED BYTE"} \>{\em Comment describing type} \\
\\
\>{\bf ASTERIX} tokens (translated when /ASTERIX is specified) :-\\
%usbyteasterix
xxxxxx\=xxxxxxxxxxxxx\=xxx\=xxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxx\kill 
\\ 
\>TOKEN     \>SUBSTITUTE     \>\>COMMENT \\
\\ 
\>$<$CONV$>$  \>\>{\tt "REAL("} \>{\em UNSIGNED BYTE to REAL conversion} \\
\>$<$EXT$>$   \>\>{\tt "ZEXT("} \>{\em UNSIGNED BYTE to REAL conversion 
extension}\\
\>$<$BRAK$>$  \>\>{\tt "))"} \>{\em Brackets for UNSIGNED BYTE to REAL 
conversion}\\
\>$<$ICONV$>$ \>\>{\tt "INT("}  \>{\em UNSIGNED BYTE to INTEGER conversion} \\
\>$<$IEXT$>$  \>\>{\tt "ZEXT("} \>{\em UNSIGNED BYTE to INTEGER conversion 
extension}\\
\>$<$IBRAK$>$   \>\>{\tt "))"}    \>{\em Brackets for UNSIGNED BYTE to INTEGER 
conversion}\\
\>$<$REV$>$     \>\>{\tt "NINT("} \>{\em REAL to UNSIGNED BYTE conversion} \\
\>$<$RVBRAK$>$  \>\>{\tt ")"}     \>{\em Brackets for REAL to UNSIGNED BYTE 
conversion} \\
\>$<$IREV$>$    \>\>{\tt " "} \>{\em INTEGER to UNSIGNED BYTE conversion} \\
\>$<$IRVBRAK$>$ \>\>{\tt " "} \>{\em Brackets for INTEGER to UNSIGNED BYTE 
conversion}\\
\\
\end{tabbing}
To produce the CHARACTER version of a program, the following
substitutions are made :-
%cHarkernal
\begin{tabbing}
xxxxxx\=xxxxxxxxxxxxx\=xxx\=xxxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxx\kill 
\>{\bf Kernel} tokens (translated by default) :-\\
\\
\>TOKEN      \>SUBSTITUTE     \>\> COMMENT \\
\\
\>$<$T$>$     \>{\tt "C"}             \>\>{\em Subroutine name extension} \\
\>$<$TYPE$>$  \>{\tt "CHARACTER*(*)"} \>\>{\em Argument type definition} \\
\>$<$LTYPE$>$ \>{\tt "CHARACTER*$<$CHASIZ$>$"} \>\>{\em Local type definition (CHASIZ = 200 by default)} \\
\>$<$CONST$>$ \>{\tt "?"}            \>\>{\em String to be appended to 
constant} \\
\>$<$HTYPE$>$ \>{\tt "\_CHAR"}       \>\>{\em HDS type} \\
\>$<$COMM$>$  \>{\tt "CHARACTER"}    \>\>{\em Comment describing type} \\
\\
\>{\bf ASTERIX} tokens (translated when /ASTERIX is specified) :- \\
%charasterix
\\
\>TOKEN      \>SUBSTITUTE     \>\> COMMENT \\
\\
\>$<$CONV$>$  \>\>{\tt "?"} \>{\em CHARACTER to REAL conversion}\\
\>$<$EXT$>$   \>\>{\tt "?"} \>{\em CHARACTER to REAL conversion extension}\\
\>$<$BRAK$>$  \>\>{\tt "?"} \>{\em Brackets for CHARACTER to REAL conversion}\\
\>$<$ICONV$>$ \>\>{\tt "?"} \>{\em CHARACTER to INTEGER conversion}\\
\>$<$IEXT$>$  \>\>{\tt "?"} \>{\em CHARACTER to INTEGER conversion extension}\\
\>$<$IBRAK$>$ \>\>{\tt "?"} \>{\em Brackets for CHARACTER to INTEGER 
conversion} \\
\>$<$REV$>$   \>\>{\tt "?"} \>{\em REAL to CHARACTER conversion} \\
\>$<$RVBRAK$>$ \>\>{\tt "?"} \>{\em Brackets for REAL to CHARACTER
conversion}\\
\>$<$IREV$>$  \>\>{\tt "?"} \>{\em INTEGER to CHARACTER conversion}\\
\>$<$IRVBRAK$>$ \>\>{\tt "?"} \>{\em Brackets for INTEGER to CHARACTER
conversion} \\
\end{tabbing}
To produce the DOUBLE PRECISION version of a program, the following
substitutions are made :-
%dpkernal
\begin{tabbing}
xxxxxx\=xxxxxxxxxxxxx\=xxx\=xxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxx\kill 
\>{\bf Kernel} tokens (translated by default) :-\\
\\
\>TOKEN     \>SUBSTITUTE  \>\>COMMENT \\
\\
\>$<$T$>$     \>{\tt "D"}               \>\>{\em Subroutine name extension}\\
\>$<$TYPE$>$  \>{\tt "DOUBLE PRECISION"}\>\>{\em Argument type definition}\\
\>$<$LTYPE$>$ \>{\tt "DOUBLE PRECISION"}\>\>{\em Local type definition}\\
\>$<$CONST$>$ \>{\tt ".0D0"}            \>\>{\em String to be appended to 
constant} \\
\>$<$HTYPE$>$ \>{\tt "\_DOUBLE"}        \>\>{\em HDS type}\\
\>$<$COMM$>$  \>{\tt "DOUBLE PRECISION"}\>\>{\em Comment describing type}\\
\end{tabbing}
\newpage                 
\begin{tabbing}
xxxxxx\=xxxxxxxxxxxxx\=xxx\=xxxxxxxxxxx\=xxxxxxxxxxxxxxxxxx\kill 
\>{\bf ASTERIX} tokens (translated when /ASTERIX is specified) :-\\
%dpasterix
\\
\>TOKEN  \>SUBSTITUTE  \>\>COMMENT \\
\\
\>$<$CONV$>$ \>\>{\tt "SNGL("} \>{\em DOUBLE PRECISION to REAL conversion}\\
\>$<$EXT$>$  \>\>{\tt " "}     \>{\em DOUBLE PRECISION to REAL conversion
extension}\\
\>$<$BRAK$>$ \>\>{\tt ")"}     \>{\em Brackets for DOUBLE PRECISION to
REAL conversion}\\
\>$<$ICONV$>$ \>\>{\tt "NINT("} \>{\em DOUBLE PRECISION to INTEGER conversion}\\
\>$<$IEXT$>$  \>\>{\tt " "}   \>{\em DOUBLE PRECISION to INTEGER conversion
extension}\\
\>$<$IBRAK$>$ \>\>{\tt ")"} \>{\em Brackets for DOUBLE PRECISION to INTEGER 
conversion}\\
\>$<$REV$>$   \>\>{\tt "DBLE("} \>{\em REAL to DOUBLE PRECISION conversion}\\
\>$<$RVBRAK$>$ \>\>{\tt ")"}    \>{\em Brackets for REAL to DOUBLE PRECISION 
conversion}\\
\>$<$IREV$>$  \>\>{\tt "DBLE("} \>{\em INTEGER to DOUBLE PRECISION
conversion}\\
\>$<$IRVBRAK$>$ \>\>{\tt ")"} \>{\em Brackets for INTEGER to DOUBLE
PRECISION conversion}\\
\end{tabbing}
To produce the INTEGER version of a program, the following substitutions
are made :-
%intkernal
\begin{tabbing}
xxxxxx\=xxxxxxxxxxxxx\=xxx\=xxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxx\kill 
\>{\bf Kernel} tokens (translated by default) :-\\
\\
\>TOKEN  \>SUBSTITUTE   \>\>COMMENT\\
\\
\>$<$T$>$     \>\>{\tt "I"}         \>{\em Subroutine name extension}\\
\>$<$TYPE$>$  \>\>{\tt "INTEGER"}   \>{\em Argument type definition}\\
\>$<$LTYPE$>$ \>\>{\tt "INTEGER"}   \>{\em Local type definition}\\
\>$<$CONST$>$ \>\>{\tt " "}         \>{\em String to be appended to constant}\\
\>$<$HTYPE$>$ \>\>{\tt "\_INTEGER"} \>{\em HDS type}\\
\>$<$COMM$>$  \>\>{\tt "INTEGER"}   \>{\em Comment describing type}\\
\\                                 
\>{\bf ASTERIX} tokens (translated when /ASTERIX is specified) :-\\
%intasterix
\\
\>TOKEN  \>SUBSTITUTE   \>\>COMMENT\\
\\
\>$<$CONV$>$     \>\>{\tt "REAL("} \>{\em INTEGER to REAL conversion}\\
\>$<$EXT$>$      \>\>{\tt " "}   \>{\em INTEGER to REAL conversion extension}\\
\>$<$BRAK$>$     \>\>{\tt ")"}     \>{\em Brackets for INTEGER to REAL 
conversion}\\
\>$<$ICONV$>$    \>\>{\tt " "}     \>{\em INTEGER to INTEGER conversion}\\
\>$<$IEXT$>$     \>\>{\tt " "}     \>{\em INTEGER to INTEGER conversion 
extension}\\
\>$<$IBRAK$>$    \>\>{\tt " "}     \>{\em Brackets for INTEGER to INTEGER 
conversion}\\
\>$<$REV$>$      \>\>{\tt "NINT("} \>{\em REAL to INTEGER conversion}\\
\>$<$RVBRAK$>$   \>\>{\tt ")"}     \>{\em Brackets for REAL to INTEGER 
conversion}\\
\>$<$IREV$>$     \>\>{\tt " "}     \>{\em INTEGER to INTEGER conversion}\\
\>$<$IRVBRAK$>$  \>\>{\tt " "}     \>{\em Brackets for INTEGER to INTEGER
conversion}\\
\end{tabbing}
To produce the LOGICAL version of a program, the following substitutions
are made :-
%logkernal
\begin{tabbing}
xxxxxx\=xxxxxxxxxxxxx\=xxx\=xxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxx\kill 
\>{\bf Kernel} tokens (translated by default) :-\\
\\
\>TOKEN      \>SUBSTITUTE      \>\>COMMENT \\
\\
\>$<$T$>$        \>\>{\tt "L"}         \>{\em Subroutine name extension}\\
\>$<$TYPE$>$     \>\>{\tt "LOGICAL"}   \>{\em Argument type definition}\\
\>$<$LTYPE$>$    \>\>{\tt "LOGICAL"}   \>{\em Local type definition}\\
\>$<$CONST$>$    \>\>{\tt " "}         \>{\em String to be appended to 
constant}\\
\>$<$HTYPE$>$    \>\>{\tt "\_LOGICAL"} \>{\em HDS type}\\
\>$<$COMM$>$     \>\>{\tt "LOGICAL"}   \>{\em Comment describing type}\\ 
\\                                            
\>{\bf ASTERIX} tokens (translated when /ASTERIX is specified) :-\\
%logasterix
\\
\>TOKEN      \>SUBSTITUTE      \>\>COMMENT\\
\\
\>$<$CONV$>$     \>\>{\tt "?"} \>{\em LOGICAL to REAL conversion}\\
\>$<$EXT$>$      \>\>{\tt "?"} \>{\em LOGICAL to REAL conversion extension}\\
\>$<$BRAK$>$     \>\>{\tt "?"} \>{\em Brackets for LOGICAL to REAL conversion}\\
\>$<$ICONV$>$    \>\>{\tt "?"} \>{\em LOGICAL to INTEGER conversion}\\
\>$<$IEXT$>$     \>\>{\tt "?"} \>{\em LOGICAL to INTEGER conversion extension}\\
\>$<$IBRAK$>$    \>\>{\tt "?"} \>{\em Brackets for LOGICAL to INTEGER 
conversion}\\
\>$<$REV$>$      \>\>{\tt "?"} \>{\em REAL to LOGICAL conversion}\\
\>$<$RVBRAK$>$   \>\>{\tt "?"} \>{\em Brackets for REAL to LOGICAL conversion}\\
\>$<$IREV$>$     \>\>{\tt "?"} \>{\em INTEGER to LOGICAL conversion}\\
\>$<$IRVBRAK$>$  \>\>{\tt "?"} \>{\em Brackets for INTEGER to LOGICAL 
conversion}\\
\end{tabbing}
To produce the REAL version of a program, the following substitutions
are made :-
%realkernal
\begin{tabbing}
xxxxxx\=xxxxxxxxxxxxx\=xxx\=xxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxx\kill 
\>{\bf Kernel} tokens (translated by default) :-\\
\\
\>TOKEN      \>SUBSTITUTE      \>\>COMMENT\\
\\
\>$<$T$>$     \>\>{\tt "R"}     \>{\em Subroutine name extension}\\
\>$<$TYPE$>$  \>\>{\tt "REAL"}  \>{\em Argument type definition}\\
\>$<$LTYPE$>$ \>\>{\tt "REAL"}  \>{\em Local type definition}\\
\>$<$CONST$>$ \>\>{\tt ".0E0"}  \>{\em String to be appended to constant}\\
\>$<$HTYPE$>$ \>\>{\tt "\_REAL"} \>{\em HDS type}\\
\>$<$COMM$>$  \>\>{\tt "REAL"}  \>{\em Comment describing type}\\
\\
\>{\bf ASTERIX} tokens (translated when /ASTERIX is specified) :-\\
%realasterix
\\
\>TOKEN      \>SUBSTITUTE      \>\>COMMENT\\
\\
\>$<$CONV$>$     \>\>{\tt " "}     \>{\em REAL to REAL conversion}\\
\>$<$EXT$>$      \>\>{\tt " "}     \>{\em REAL to REAL conversion extension}\\
\>$<$BRAK$>$     \>\>{\tt " "}  \>{\em Brackets for REAL to REAL conversion}\\
\>$<$ICONV$>$    \>\>{\tt "NINT("} \>{\em REAL to INTEGER conversion}\\
\>$<$IEXT$>$     \>\>{\tt " "}  \>{\em REAL to INTEGER conversion extension}\\
\>$<$IBRAK$>$    \>\>{\tt ")"} \>{\em Brackets for REAL to INTEGER conversion}\\
\>$<$REV$>$      \>\>{\tt " "}     \>{\em REAL to REAL conversion}\\
\>$<$RVBRAK$>$   \>\>{\tt " "}   \>{\em Brackets for REAL to REAL conversion}\\
\>$<$IREV$>$     \>\>{\tt "REAL("} \>{\em INTEGER to REAL conversion}\\
\>$<$IRVBRAK$>$  \>\>{\tt ")"} \>{\em Brackets for INTEGER to REAL conversion}\\
\end{tabbing}           
To produce the WORD version of a program, the following substitutions
are made :-
%wordkernal
\begin{tabbing}
xxxxxx\=xxxxxxxxxxxxx\=xxx\=xxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxx\kill 
\>{\bf Kernel} tokens (translated by default) :-\\
\\
\>TOKEN      \>SUBSTITUTE     \>\>COMMENT\\
\\
\>$<$T$>$     \>\>{\tt "W"}         \>{\em Subroutine name extension}\\
\>$<$TYPE$>$  \>\>{\tt "INTEGER*2"} \>{\em Argument type definition}\\
\>$<$LTYPE$>$ \>\>{\tt "INTEGER*2"} \>{\em Local type definition}\\
\>$<$CONST$>$ \>\>{\tt " "}         \>{\em String to be appended to constant}\\
\>$<$HTYPE$>$ \>\>{\tt "\_WORD"}    \>{\em HDS type}\\
\>$<$COMM$>$  \>\>{\tt "WORD"}      \>{\em Comment describing type}\\
\\                                         
\>ASTERIX tokens (translated when /ASTERIX is specified) :-\\
%wordasterix
\\
\>TOKEN      \>SUBSTITUTE      \>\>COMMENT\\
\\
\>$<$CONV$>$     \>\>{\tt "REAL("} \>{\em WORD to REAL conversion}\\
\>$<$EXT$>$      \>\>{\tt " "}     \>{\em WORD to REAL conversion extension}\\
\>$<$BRAK$>$     \>\>{\tt ")"}   \>{\em Brackets for WORD to REAL conversion}\\
\>$<$ICONV$>$    \>\>{\tt " "}     \>{\em WORD to INTEGER conversion}\\
\>$<$IEXT$>$     \>\>{\tt " "}   \>{\em WORD to INTEGER conversion extension}\\
\>$<$IBRAK$>$    \>\>{\tt " "} \>{\em Brackets for WORD to INTEGER conversion}\\
\>$<$REV$>$      \>\>{\tt "NINT("} \>{\em REAL to WORD conversion}\\
\>$<$RVBRAK$>$   \>\>{\tt ")"}  \>{\em Brackets for REAL to WORD conversion}\\
\>$<$IREV$>$     \>\>{\tt " "}     \>{\em INTEGER to WORD conversion}\\
\>$<$IRVBRAK$>$  \>\>{\tt " "} \>{\em Brackets for INTEGER to WORD conversion}\\
\end{tabbing}                             
To produce the UNSIGNED WORD version of a program, the following
substitutions are made :-
%uswordkernal
\begin{tabbing}
xxxxxx\=xxxxxxxxxxxxx\=xxx\=xxxxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxx\kill 
\>{\bf Kernel} tokens (translated by default) :-\\
\\
\>TOKEN      \>SUBSTITUTE     \>\>COMMENT\\
\\
\>$<$T$>$     \>\>{\tt "UW"}            \>{\em Subroutine name extension}\\
\>$<$TYPE$>$  \>\>{\tt "INTEGER*2"}     \>{\em Argument type definition}\\
\>$<$LTYPE$>$ \>\>{\tt "INTEGER*2"}     \>{\em Local type definition}\\
\>$<$CONST$>$ \>\>{\tt " "}         \>{\em String to be appended to constant}\\
\>$<$HTYPE$>$ \>\>{\tt "\_UWORD"}       \>{\em HDS type}\\
\>$<$COMM$>$  \>\>{\tt "UNSIGNED WORD"} \>{\em Comment describing type}\\
\\              
\>{\bf ASTERIX} tokens (translated when /ASTERIX is specified) :-\\
%uswordasterix
xxxxxx\=xxxxxxxxxxxxx\=xxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxx\kill 
\\
\>TOKEN      \>SUBSTITUTE      \>\>COMMENT\\
\\
\>$<$CONV$>$     \>\>{\tt "REAL("} \>{\em UNSIGNED WORD to REAL conversion}\\
\>$<$EXT$>$      \>\>{\tt "ZEXT("} \>{\em UNSIGNED WORD to REAL conversion 
extension}\\
\>$<$BRAK$>$     \>\>{\tt "))"}    \>{\em Brackets for UNSIGNED WORD to REAL 
conversion}\\
\>$<$ICONV$>$    \>\>{\tt " "}     \>{\em UNSIGNED WORD to INTEGER conversion}\\
\>$<$IEXT$>$     \>\>{\tt "ZEXT("} \>{\em UNSIGNED WORD to INTEGER conversion 
extension}\\
\>$<$IBRAK$>$    \>\>{\tt ")"}     \>{\em Brackets for UNSIGNED WORD to INTEGER
 conversion}\\
\>$<$REV$>$      \>\>{\tt "NINT("} \>{\em REAL to UNSIGNED WORD conversion}\\
\>$<$RVBRAK$>$   \>\>{\tt ")"}     \>{\em Brackets for REAL to UNSIGNED WORD 
conversion}\\
\>$<$IREV$>$     \>\>{\tt " "}     \>{\em INTEGER to UNSIGNED WORD conversion}\\
\>$<$IRVBRAK$>$  \>\>{\tt " "}     \>{\em Brackets for INTEGER to UNSIGNED WORD
 conversion}\\
\end{tabbing}                             
\newpage
\begin{center}
{\Large\bf APPENDIX 2}

{\large\bf A FULL LIST OF CONVERSION FUNCTIONS}
\end{center}

This appendix gives the conversion functions which are available using
the DCV utility, and describes which ones are statement functions and
which ones are external. 

The following conversions are defined as in-line statement functions in 
`DCV\_FUN'. (They are either trivial, or are mentioned in the Fortran
manual as legal conversions) :-

\begin{tabbing}
xxxx\=xxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\=xxx\kill 
\>{\bf Function name}   \>{\bf Conversion}\\
\\
\>DCV\_BTOB   \>BYTE to BYTE \\
\>DCV\_UBTOUB \>UNSIGNED BYTE to UNSIGNED BYTE \\
\>DCV\_UBTOD  \>UNSIGNED BYTE to DOUBLE PRECISION \\
\>DCV\_DTOD   \>DOUBLE PRECISION to DOUBLE PRECISION \\
\>DCV\_ITOD   \>INTEGER to DOUBLE PRECISION \\
\>DCV\_RTOD   \>REAL to DOUBLE PRECISION \\
\>DCV\_WTOD   \>WORD to DOUBLE PRECISION \\
\>DCV\_UWTOD  \>UNSIGNED WORD to DOUBLE PRECISION \\
\>DCV\_UBTOI  \>UNSIGNED BYTE to INTEGER \\
\>DCV\_DTOI   \>DOUBLE PRECISION to INTEGER \\
\>DCV\_ITOI   \>INTEGER to INTEGER \\
\>DCV\_RTOI   \>REAL to INTEGER \\
\>DCV\_UWTOI  \>UNSIGNED WORD to INTEGER \\
\>DCV\_UBTOR  \>UNSIGNED BYTE to REAL \\
\>DCV\_DTOR   \>DOUBLE PRECISION to REAL \\
\>DCV\_ITOR   \>INTEGER to REAL \\
\>DCV\_RTOR   \>REAL to REAL \\
\>DCV\_WTOR   \>WORD to REAL \\
\>DCV\_UWTOR  \>UNSIGNED WORD to REAL \\
\>DCV\_UBTOW  \>UNSIGNED BYTE to WORD \\
\>DCV\_DTOW   \>DOUBLE PRECISION to WORD       \>*** \\
\>DCV\_RTOW   \>REAL to WORD                   \>*** \\
\>DCV\_WTOW   \>WORD to WORD \\
\>DCV\_UBTOUW \>UNSIGNED BYTE to UNSIGNED WORD \\
\>DCV\_UWTOUW \>UNSIGNED WORD to UNSIGNED WORD \\
\end{tabbing} 
Overflow errors are possible from the conversions marked ``***''

\newpage

The following conversions are also defined as in-line statement functions in
`DCV\_FUN'. However, they are not mentioned in the Fortran manual and should be
used with caution. Trial and error has shown that the Fortran compliers on all
Starlink supported systems will make the conversions successfully :-

\begin{tabbing}
xxxx\=xxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\=xxx\kill 
\>{\bf Function name}      \>{\bf Conversion}\\
\\
\>DCV\_UBTOB \>UNSIGNED BYTE to BYTE          \>*** \\
\>DCV\_DTOB  \>DOUBLE PRECISION to BYTE       \>*** \\
\>DCV\_ITOB  \>INTEGER to BYTE                \>*** \\
\>DCV\_RTOB  \>REAL to BYTE                   \>*** \\
\>DCV\_WTOB  \>WORD to BYTE                   \>*** \\
\>DCV\_UWTOB \>UNSIGNED WORD to BYTE          \>*** \\
\>DCV\_BTOD  \>BYTE to DOUBLE PRECISION \\
\>DCV\_BTOI  \>BYTE to INTEGER \\
\>DCV\_WTOI  \>WORD to INTEGER \\
\>DCV\_BTOR  \>BYTE to REAL \\
\>DCV\_BTOW  \>BYTE to WORD \\
\>DCV\_ITOW  \>INTEGER to WORD                \>*** \\
\end{tabbing}
Overflow errors are possible from the conversions marked ``***''

The following conversions are defined as external
functions in the DCV object library :-
\begin{tabbing}
xxxx\=xxxxxxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\=xxx\kill 
\>{\bf Function name}     \>{\bf Conversion}\\
\\
\>DCV\_BTOUB  \>BYTE to UNSIGNED BYTE \\
\>DCV\_DTOUB  \>DOUBLE PRECISION to UNSIGNED BYTE \\
\>DCV\_ITOUB  \>INTEGER to UNSIGNED BYTE \\
\>DCV\_RTOUB  \>REAL to UNSIGNED BYTE \\
\>DCV\_WTOUB  \>WORD to UNSIGNED BYTE \\
\>DCV\_UWTOUB \>UNSIGNED WORD to UNSIGNED BYTE \\
\>DCV\_BTOUW  \>BYTE to UNSIGNED WORD \\
\>DCV\_UBTOUW \>UNSIGNED BYTE to UNSIGNED WORD \\
\>DCV\_DTOUW  \>DOUBLE PRECISION to UNSIGNED WORD \\
\>DCV\_ITOUW  \>INTEGER to UNSIGNED WORD \\
\>DCV\_RTOUW  \>REAL to UNSIGNED WORD \\
\>DCV\_WTOUW  \>WORD to UNSIGNED WORD \\
\end{tabbing} 
Overflow errors are possible in ALL these conversions, so range checking is 
essential.

\end{document}
