\input SUN93LAY

\pageno=7
\nopagenumbers
\parindent=0in
\headline={\ifnum\pageno=1
            \else\it SUN/93.2\hss \rm\folio\fi}

{\bf Appendix. Font Selection}

Variations of font type can be achieved with the control
sequences $\backslash$rm,$\backslash$it,$\backslash$sl,$\backslash$tt {\it etc$\ldotp$} 
Similar commands can be defined using the $\backslash$font command to access 
any available fonts. For example  a command $\backslash$inch
to select the `CMINCH' font (Computer Modern INCH high letters) can be defined thus:
\smallskip
\tt
\settabs\+\indent&slashfont=CMB10spacescaled 1000&                          \cr

\+&$\backslash$font$\backslash$inch=CMINCH &\% inch high letters\cr
\rm
\noindent Subsequent use of the control sequence $\backslash$inch
is identical to the use of $\backslash$rm, $\backslash$it {\it etc.}


In addition to its normal size, each font is provided at a range of
magnifications in a geometric progression. 
Specifically, each font is available at magnifications of 1, 1.2, 
1.2$\times$1.2, {\it etc.} Altogether in the current Starlink version
of {\TeX}, magnifications of 1, 1.2, 1.44, 1.728, 2.074, 2.488, 
2.986, 3.583 and 4.3 are available.
(The first five of these numbers correspond to $\backslash$magstep0, 
$\backslash$magstep1, $\backslash$magstep2, $\backslash$magstep3,
$\backslash$magstep4 and $\backslash$magstep5 respectively.
Additionally there is an intermediate size between 1 and 1.2, 
$\sqrt{1.2}=1.095$ --- equivalent to $\backslash$magstephalf.)

Magnified fonts can be selected using a `scaling factor' which is 
1000$\times$the required magnification.
For example, various sizes of the font `cmb10' can be accessed by
defining the control sequences $\backslash$fzero, $\backslash$fhalf, 
$\backslash$fone, $\backslash$ftwo as follows:	
\smallskip
\tt
\settabs\+\indent&slashfont=CMB10spacescaled 1000&                          \cr

\+&$\backslash$font$\backslash$fzero=CMB10  
&\%  unscaled cmb10\cr
\+&$\backslash$font$\backslash$fhalf=CMB10 scaled 1095 &\% 1.095$\times$normal size 
(equivalent to $\backslash$magstephalf)\cr
\+&$\backslash$font$\backslash$fone=CMB10 scaled 1200 &\% 1.2$\times$normal size 
(equivalent to $\backslash$magstep1)\cr
\+&$\backslash$font$\backslash$ftwo=CMB10 scaled 1440 &\% 1.44$\times$normal size
(equivalent to $\backslash$magstep2)\cr
\rm
%Thus you can produce a sentence like this

\TeX\ has a large but limited capacity for fonts. For example, this appendix 
could not be generated in a single \TeX\ run.
Additionally there is  a limit to the number of fonts the
CANON laser printer can have on a page. 
However these restrictions will only inconvenience
those who require more than about thirty fonts in a document.

The fonts available in the current Starlink version of \TeX\ are shown below,
and the maximum scaling factor (MSF) at which each is available is indicated.
It can be assumed those sizes smaller than the maximum are present, thus 
an entry of 4300
indicates that a font is available at scaling factors 1000, 1095, 
1200, 1440, 1728, 2074, 2488, 2986, 3583 and 4300.
Attempts to use a non-existent font-size will not cause errors during {\TeX}ing
but will be reported at the stage when  device-dependent output is 
being generated.
\bigskip


\settabs\+\indent&ABCDEFGHIJKLMNOPQRSTUVWXYZabcdefghijklmnopqrstuvwxyz&
CMSSI101234567 &1440~~\cr\medskip
\+&{\bf Font appearance}&{\bf Font name}&{\bf MSF}\cr
\bigskip

%\font\joa=CIRCLE10 scaled\magstep0
%\font\job=CIRCLEW10 scaled\magstep0
\font\joc=CMB10 scaled\magstep0
\font\jod=CMBSY10 scaled\magstep0
\font\joe=CMBX10 scaled\magstep0
\font\jof=CMBX12 scaled\magstep0
\font\jog=CMBX5 scaled\magstep0
\font\joh=CMBX6 scaled\magstep0
\font\joi=CMBX7 scaled\magstep0
\font\joj=CMBX8 scaled\magstep0
\font\jok=CMBX9 scaled\magstep0
\font\jol=CMBXSL10 scaled\magstep0
\font\jom=CMBXTI10 scaled\magstep0
\font\jon=CMCSC10 scaled\magstep0
\font\joo=CMDUNH10 scaled\magstep0
\font\jop=CMEX10 scaled\magstep0
\font\joq=CMFF10 scaled\magstep0
%%%%%%%\font\jor=CMFI10 scaled\magstep0
\font\jos=CMFIB8 scaled\magstep0
\font\jot=CMINCH scaled\magstep0
%\font\jou=CMIT10 scaled\magstep0   % originally cmitt10
\font\jov=CMMI10 scaled\magstep0
\font\jow=CMMI12 scaled\magstep0
\font\jox=CMMI5 scaled\magstep0
\font\joy=CMMI6 scaled\magstep0
\font\joz=CMMI7 scaled\magstep0
\font\joaa=CMMI8 scaled\magstep0
\font\jobb=CMMI9 scaled\magstep0
\font\jocc=CMMIB10 scaled\magstep0
\font\jodd=CMR10 scaled\magstep0
\font\joee=CMR12 scaled\magstep0
\font\joff=CMR17 scaled\magstep0
\font\jogg=CMR5 scaled\magstep0
\font\johh=CMR6 scaled\magstep0
\font\joii=CMR7 scaled\magstep0
\font\jojj=CMR8 scaled\magstep0
\font\jokk=CMR9 scaled\magstep0
%\font\jokka=CMSA10 scaled\magstep0  % EXTRA ?
\font\joll=CMSL10 scaled\magstep0
\font\jomm=CMSL12 scaled\magstep0

%\+&{\joa ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CIRCLE10 }& 1440\cr\medskip
%\+&{\job ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CIRCLEW10 }& 1440\cr\medskip
\+&{\joc ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMB10 }& 1440\cr\medskip
\+&{\jod ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMBSY10 }& 1440\cr\medskip
\+&{\joe ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMBX10 }& 2986\cr\medskip
\+&{\jof ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMBX12 }& 1440\cr\medskip
\+&{\jog ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMBX5 }& 2986\cr\medskip
\+&{\joh ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMBX6 }& 1440\cr\medskip
\+&{\joi ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMBX7 }& 2986\cr\medskip

\vfill\eject
\+&{\bf Font appearance}&{\bf Font name}&{\bf MSF}\cr
\bigskip

\+&{\joj ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMBX8 }& 1440\cr\medskip
\+&{\jok ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMBX9 }& 1440\cr\medskip
\+&{\jol ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMBXSL10 }& 1440\cr\medskip
\+&{\jom ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMBXTI10 }& 1440\cr\medskip
\+&{\jon ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMCSC10 }& 1728\cr\medskip
\+&{\joo ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMDUNH10 }& 1440\cr\medskip
\+&{\jop ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMEX10 }& 1728\cr\medskip
\+&{\joq ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMFF10 }& 1440\cr\medskip
%\+&{\jor ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMFI10 }& 1440\cr\medskip
\+&{\jos ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMFIB8 }& 1440\cr\medskip
\+&{\jot ABC  }& {  CMINCH }& 1440\cr\medskip
%\+&{\jou ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMIT10 }&1440\cr\medskip %?
\+&{\jov ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMMI10 }& 2986\cr\medskip
\+&{\jow ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMMI12 }& 1440\cr\medskip
\+&{\jox ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMMI5 }& 2986\cr\medskip
\+&{\joy ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMMI6 }& 1440\cr\medskip
\+&{\joz ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMMI7 }& 2986\cr\medskip
\+&{\joaa ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMMI8 }& 4300\cr\medskip
\+&{\jobb ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMMI9 }& 1440\cr\medskip
\+&{\jocc ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMMIB10 }& 1440\cr\medskip
\+&{\jodd ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMR10 }& 2986\cr\medskip
\+&{\joee ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMR12 }& 1440\cr\medskip
\+&{\joff ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMR17 }& 1440\cr\medskip
\+&{\jogg ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMR5 }& 2986\cr\medskip
\+&{\johh ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMR6 }& 1440\cr\medskip
\+&{\joii ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMR7 }& 2986\cr\medskip
\+&{\jojj ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMR8 }& 1440\cr\medskip
\+&{\jokk ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMR9 }& 1440\cr\medskip
\+&{\joll ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMSL10 }& 2986\cr\medskip
\+&{\jomm ABCDEFGHIJKLMNOPQRSTUVWXYZ}& {  CMSL12 }& 1440\cr\medskip
\bye
